---
title: "Results - revision"
author: "Martin Bulla"
date: "`r Sys.time()`"
output:
    html_document:
        toc: true
        toc_float: true
        toc_depth: 5
        code_folding: hide
        link-citations: yes
---

```{r r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, cache = FALSE)
```

###### Code to load tools and data

```{r start, echo = T, results = 'hide', warning=FALSE}
 # packages
    require(arm)
    require(data.table)
    require(effects)
    require(foreach)
    require(here)
    require(ggimage)
    require(ggplot2)
    require(ggpubr)
    require(ggsci)
    require(ggtext)
    require(grid)
    require(gtable)
    require(MASS)
    require(multcomp)
    require(optimx)
    require(performance)  
    require(PerformanceAnalytics)
    require(png)
    require(RColorBrewer)
    require(rmeta)
    require(rphylopic)
    require(scales)
    require(viridis)
 # constants
    round_ = 3 # number of decimal places to round model coefficients
    nsim = 5000 # number of simulations to extract estimates and 95%CrI
    ax_lines = "grey60" # defines color of the axis lines
    #colors <- c("#999999", "#E69F00", "#56B4E9") #viridis(3)
    set.seed(42)
    #width_ = .7 # spacing between error bars
    #col_ = c(brewer.pal(n =12, name = "Paired"), 'grey30','grey80')
 # functions
    # to add images to panels
      annotation_custom2 <- function (grob, xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, data) {
        layer(data = data, stat = StatIdentity, position = PositionIdentity, 
            geom = ggplot2:::GeomCustomAnn,
            inherit.aes = TRUE, params = list(grob = grob, 
                                              xmin = xmin, xmax = xmax, 
                                              ymin = ymin, ymax = ymax))
      }
    # to remove ggplot components
        gtable_filter_remove <- function (x, name, trim = TRUE){
          matches <- !(x$layout$name %in% name)
          x$layout <- x$layout[matches, , drop = FALSE]
          x$grobs <- x$grobs[matches]
          if (trim) 
            x <- gtable_trim(x)
          x
        }
    # customized ggplot theme
        theme_MB = theme(  
                  title = element_text(size=8, colour="grey30"),
                  axis.line = element_blank(),
                  #axis.line = element_line(colour="grey70", size=0.25),
                  axis.title = element_text(size=7, colour="grey30"),
                  axis.title.y = element_text(vjust=3.5),
                  axis.title.x = element_text(vjust=1),
                  axis.text = element_text(size=6),#, vjust = 0.5, hjust=1),# margin=units(0.5,"mm")),
                  axis.ticks.length=unit(0.5,"mm"),
                  axis.ticks = element_line(colour = "grey70", size = 0.1),
                  #axis.ticks.margin,
                  
                  strip.text.x = element_text(size = 6, color="grey30",  margin=margin(1,1,1,1,"mm")), #grey50
                  strip.text.y = element_text(size = 6, color="grey30",  margin=margin(1,1,1,1,"mm")), #grey50
                  strip.background = element_rect(fill="grey99",colour="grey70", size=0.25),
                    #strip.background = element_blank(), 
                    #strip.text = element_blank(),
                  panel.spacing = unit(0, "mm"),
                  panel.background=element_blank(),
                  panel.border = element_rect(colour="grey70", size=0.1, fill = NA), #panel.border=element_blank(),
                  panel.grid = element_blank(),

                  legend.text=element_text(size=6),
                  legend.title=element_text(size=6),
                  legend.key = element_rect(colour = NA, fill = NA),
                  legend.key.height= unit(0.5,"line"),
                  legend.key.width = unit(0.25, "cm"),
                  legend.margin = margin(0,0,0,0, unit="cm"),
                  legend.box.margin = margin(l = -6), #legend.justification = c(-1,0),
                  legend.background = element_blank()
                  )  
    # for estimates
        est_out =function(model = m, label = "", nsim = 5000){
            bsim = sim(model, n.sim=5000) 
            v = apply(bsim@fixef, 2, quantile, prob=c(0.5))
            ci = apply(bsim@fixef, 2, quantile, prob=c(0.025,0.975)) 
            sd = apply(bsim@fixef, 2, sd)
            data.table(predictor=rownames(coef(summary(model))),estimate=v, lwr=ci[1,], upr=ci[2,], sd = sd, model = paste(label, "N =", nobs(model)))
          }
    # change color
      change_col = function(replace_black, theimg) {
        r_b = col2rgb(replace_black) / 255
        #theimg[theimg == 1] <- 2
        for (i in 1:3) {
            theimg[,,i][theimg[,,i] == 0] <- r_b[i]
        }
        return(theimg)
        }
    # for Supplementary Table output based on sim
      m_out = function(model = m, type = "mixed", 
        name = "define", dep = "define", fam = 'Gaussian',
        round_ = 3, nsim = 5000, aic = FALSE, save_sim = 'Data/model_sim/', back_tran = FALSE, perc_ = 1){
          # perc_ 1 = proportion or 100%
        bsim = sim(model, n.sim=nsim)  
        
        if(save_sim!=FALSE){save(bsim, file = paste0(save_sim, name,'.RData'))}
       
        if(type != "mixed"){
          v = apply(bsim@coef, 2, quantile, prob=c(0.5))
          ci = apply(bsim@coef, 2, quantile, prob=c(0.025,0.975)) 

          if(back_tran == TRUE & fam == "binomial"){
           v = perc_*plogis(v)
           ci = perc_*plogis(ci)
           }
          if(back_tran == TRUE & fam == "binomial_logExp"){
                v = perc_*(1-plogis(v))
                ci = perc_*(1-plogis(ci))
                ci = rbind(ci[2,],ci[1,])
               }

          if(back_tran == TRUE & fam == "Poisson"){
           v = exp(v)
           ci = exp(ci)
          }

         oi=data.frame(type='fixed',effect=rownames(coef(summary(model))),estimate=v, lwr=ci[1,], upr=ci[2,])
          rownames(oi) = NULL
          oi$estimate_r=round(oi$estimate,round_)
          oi$lwr_r=round(oi$lwr,round_)
          oi$upr_r=round(oi$upr,round_)
          if(perc_ == 100){
           oi$estimate_r = paste0(oi$estimate_r,"%")
           oi$lwr_r = paste0(oi$lwr_r,"%")
           oi$upr_r = paste0(oi$upr_r,"%")
          }
         x=data.table(oi[c('type',"effect", "estimate_r","lwr_r",'upr_r')]) 
       
        }else{
         v = apply(bsim@fixef, 2, quantile, prob=c(0.5))
         ci = apply(bsim@fixef, 2, quantile, prob=c(0.025,0.975)) 

         if(back_tran == TRUE & fam == "binomial"){
          v = perc_*plogis(v)
          ci = perc_*plogis(ci)
         }
          if(back_tran == TRUE & fam == "binomial_logExp"){
                v = perc_*(1-plogis(v))
                ci = perc_*(1-plogis(ci))
                ci = rbind(ci[2,],ci[1,])
               }

          if(back_tran == TRUE & fam == "Poisson"){
            v = exp(v)
            ci = exp(ci)
         }

        oi=data.table(type='fixed',effect=rownames(coef(summary(model))),estimate=v, lwr=ci[1,], upr=ci[2,])
            rownames(oi) = NULL
            oi[,estimate_r := round(estimate,round_)]
            oi[,lwr_r := round(lwr,round_)]
            oi[,upr_r :=round(upr,round_)]
            if(perc_ == 100){
             oi[,estimate_r := paste0(estimate_r,"%")]
             oi[,lwr_r := paste0(lwr_r,"%")]
             oi[,upr_r := paste0(upr_r,"%")]
            }
         oii=oi[,c('type',"effect", "estimate_r","lwr_r",'upr_r')] 
        
         l=data.frame(summary(model)$varcor)
         l = l[is.na(l$var2),]
         l$var1 = ifelse(is.na(l$var1),"",l$var1)
         l$pred = paste(l$grp,l$var1)

         q050={}
         q025={}
         q975={}
         pred={}
         
         # variance of random effects
         for (ran in names(bsim@ranef)) {
           ran_type = l$var1[l$grp == ran]
           for(i in ran_type){
            q050=c(q050,quantile(apply(bsim@ranef[[ran]][,,ran_type], 1, var), prob=c(0.5)))
            q025=c(q025,quantile(apply(bsim@ranef[[ran]][,,ran_type], 1, var), prob=c(0.025)))
            q975=c(q975,quantile(apply(bsim@ranef[[ran]][,,ran_type], 1, var), prob=c(0.975)))
            pred= c(pred,paste(ran, i))
            }
           }
         # residual variance
         q050=c(q050,quantile(bsim@sigma^2, prob=c(0.5)))
         q025=c(q025,quantile(bsim@sigma^2, prob=c(0.025)))
         q975=c(q975,quantile(bsim@sigma^2, prob=c(0.975)))
         pred= c(pred,'Residual')

         ri=data.table(type='random %',effect=pred, estimate_r=round(100*q050/sum(q050)), lwr_r=round(100*q025/sum(q025)), upr_r=round(100*q975/sum(q975)))
           
         ri[lwr_r>upr_r, lwr_rt := upr_r]
         ri[lwr_r>upr_r, upr_rt := lwr_r]
         ri[!is.na(lwr_rt), lwr_r := lwr_rt]
         ri[!is.na(upr_rt), upr_r := upr_rt]
         ri$lwr_rt = ri$upr_rt = NULL

         ri[,estimate_r := paste0(estimate_r,'%')]
         ri[,lwr_r := paste0(lwr_r,'%')]
         ri[,upr_r := paste0(upr_r,'%')]
        
        x = data.table(rbind(oii,ri))
        }
        
        x[1, model := name]                                                                
        x[1, response := dep]                                                                
        x[1, error_structure := fam]      
        N = length(resid(model))                                                          
        x[1, N := N ]                                                                

        x=x[ , c('model', 'response', 'error_structure', 'N', 'type',"effect", "estimate_r","lwr_r",'upr_r')] 

        if (aic == TRUE){   
            x[1, AIC := AIC(update(model,REML = FALSE))] 
            }
        if (aic == "AICc"){
            aicc = AICc(model)
            x[1, AICc := aicc] 
        }
        if(type == "mixed" & nrow(x[type=='random %' & estimate_r =='0%'])==0){
          R2_mar = as.numeric(r2_nakagawa(model)$R2_marginal)
          R2_con = as.numeric(r2_nakagawa(model)$R2_conditional)
          x[1, R2_mar := R2_mar]
          x[1, R2_con := R2_con]
         }
        x[is.na(x)] = ""
        return(x)
      } 
    # model assumption function
      m_ass = function(name = 'define', mo = m0, dat = d, fixed = NULL, categ = NULL, trans = "none", spatial = TRUE, temporal = TRUE, PNG = TRUE, outdir = 'outdir', n_col=8, width_ = 10, height_ = 5.5){
       l=data.frame(summary(mo)$varcor)
       l = l[is.na(l$var2),]
       nt = if(temporal==TRUE){1}else{0}
       ns = if(spatial==TRUE){7}else{0}
       n = 3+nrow(l)-1+length(fixed)+length(categ) +  nt +  ns
     
       if(PNG == TRUE){
        png(paste(outdir,name, ".png", sep=""), width=width_,height=height_,units="in",res=600) # width = 6
        par(mfrow=c(4, n_col),tcl = -0.08, cex = 0.5, cex.main = 0.9,#ceiling(n/n_col),n_col)
            oma = c(1,1,2,1),
            mar = c(2, 2, 2, 1), mgp=c(1,0,0)
            )
         }else{
          dev.new(width=width_,height=height_)
          par(mfrow=c(4,n_col), tcl = -0.08, cex = 0.5, cex.main = 0.9,#ceiling(n/n_col),n_col)
            oma = c(1,1,2,1),
            mar = c(2, 2, 2, 1), mgp=c(1,0,0)
            )
        }
       
       scatter.smooth(fitted(mo),resid(mo),col='grey');abline(h=0, lty=2, col ='red')
       scatter.smooth(fitted(mo),sqrt(abs(resid(mo))), col='grey')
       qqnorm(resid(mo), main=list("Normal Q-Q Plot: residuals"),col='grey');qqline(resid(mo), col = 'red')
       #unique(l$grp[l$grp!="Residual"])
       for(i in unique(l$grp[l$grp!="Residual"])){
        #i = "mean_year"
        ll=ranef(mo)[names(ranef(mo))==i][[1]]
        if(ncol(ll)==1){
         qqnorm(ll[,1], main = paste(i,names(ll)[1]),col='grey',);qqline(ll[,1], col ='red')
         }else{
          qqnorm(ll[,1], main = paste(i,names(ll)[1]),col='grey');qqline(ll[,1], col ='red')
          qqnorm(ll[,2], main = paste(i,names(ll)[2]),col='grey');qqline(ll[,2], col ='red')
         }
        }
        
       # variables
         scatter={} 
         for (i in rownames(summary(mo)$coef)) {
              #i = "lat_abs"
            j=sub("\\).*", "", sub(".*\\(", "",i)) 
            scatter[length(scatter)+1]=j
          }
          x = data.frame(scatter=unique(scatter)[2:length(unique(scatter))],
                          log_ = grepl("log",rownames(summary(mo)$coef)[2:length(unique(scatter))]), stringsAsFactors = FALSE)
          for (i in 1:length(fixed)){
              jj =fixed[i]
              variable=dat[, ..jj][[1]]
              if(trans[i]=='log'){
              scatter.smooth(resid(mo)~log(variable),xlab=paste('log(',jj,')',sep=''), col = 'grey');abline(h=0, lwd=1, lty = 2, col ='red')
              }else if(trans[i]=='abs'){
              scatter.smooth(resid(mo)~abs(variable),xlab=paste('abs(',jj,')',sep=''), col = 'grey');abline(h=0, lwd=1, lty = 2, col ='red')
              }else if(trans[i]=='sin'){scatter.smooth(resid(mo)~sin(variable),xlab=paste('sin(',jj,')',sep=''), col = 'grey');abline(h=0, lwd=1, lty = 2, col ='red')
              }else if(trans[i]=='cos'){scatter.smooth(resid(mo)~cos(variable),xlab=paste('cos(',jj,')',sep=''), col = 'grey');abline(h=0, lwd=1, lty = 2, col ='red')
              }else{
              scatter.smooth(resid(mo)~variable,xlab=jj,col = 'grey');abline(h=0, lwd=1, lty = 2, col ='red')
            }
           }
          
          if(length(categ)>0){
            for(i in categ){
               variable=dat[, ..i][[1]]
                boxplot(resid(mo)~variable, medcol='grey', whiskcol='grey', staplecol='grey', boxcol='grey', outcol='grey', xlab = i);abline(h=0, lty=3, lwd=1, col = 'red')
               }
          }     
              
        if(temporal == TRUE){
            acf(resid(mo), type="p", main=list("Temporal autocorrelation:\npartial series residual"))
            }
        if(spatial == TRUE){    
          spdata=data.frame(resid=resid(mo), x=dat$Lon, y=dat$Lat)
            spdata$col=ifelse(spdata$resid<0,rgb(83,95,124,100, maxColorValue = 255),ifelse(spdata$resid>0,rgb(253,184,19,100, maxColorValue = 255), 'red'))
            #cex_=c(1,2,3,3.5,4)
            cex_=c(1,1.5,2,2.5,3)
            spdata$cex=as.character(cut(abs(spdata$resid), 5, labels=cex_))
          plot(spdata$x, spdata$y,col=spdata$col, cex=as.numeric(spdata$cex), pch= 16, main=list('Spatial distribution of residuals', cex=0.8), xlab = 'Longitude', ylab = 'Latitude')
            legend("topright", pch=16, legend=c('>0','<0'), ,col=c(rgb(83,95,124,100, maxColorValue = 255),rgb(253,184,19,100, maxColorValue = 255)))
          plot(spdata$x[spdata$resid<0], spdata$y[spdata$resid<0],col=spdata$col[spdata$resid<0], cex=as.numeric(spdata$cex[spdata$resid<0]), pch= 16, main=list('residuals <0'), xlab = 'Longitude', ylab = 'Latitude')
          plot(spdata$x[spdata$resid>=0], spdata$y[spdata$resid>=0],col=spdata$col[spdata$resid>=0], cex=as.numeric(spdata$cex[spdata$resid>=0]), pch= 16, main=list('residual >=0'), xlab = 'Longitude', ylab = 'Latitude')


          # EU
          dat$res = resid(mo)
          spdata=data.frame(resid = dat$res[dat$Country!='Australia'], x=dat$Lon[dat$Country!='Australia'], y=dat$Lat[dat$Country!='Australia'])
          spdata$col=ifelse(spdata$resid<0,rgb(83,95,124,100, maxColorValue = 255),ifelse(spdata$resid>0,rgb(253,184,19,100, maxColorValue = 255), 'red'))
            #cex_=c(1,2,3,3.5,4)
            cex_=c(1,1.5,2,2.5,3)
            spdata$cex=as.character(cut(abs(spdata$resid), 5, labels=cex_))
          plot(spdata$x[spdata$resid<0], spdata$y[spdata$resid<0],col=spdata$col[spdata$resid<0], cex=as.numeric(spdata$cex[spdata$resid<0]), pch= 16, main=list('EU -  residuals <0'), xlab = 'Longitude', ylab = 'Latitude')
          plot(spdata$x[spdata$resid>=0], spdata$y[spdata$resid>=0],col=spdata$col[spdata$resid>=0], cex=as.numeric(spdata$cex[spdata$resid>=0]), pch= 16, main=list('EU residuals >=0)'), xlab = 'Longitude', ylab = 'Latitude')

          # Australia
          spdata=data.frame(resid = dat$res[dat$Country=='Australia'], x=dat$Lon[dat$Country=='Australia'], y=dat$Lat[dat$Country=='Australia'])
          spdata$col=ifelse(spdata$resid<0,rgb(83,95,124,100, maxColorValue = 255),ifelse(spdata$resid>0,rgb(253,184,19,100, maxColorValue = 255), 'red'))
            #cex_=c(1,2,3,3.5,4)
            cex_=c(1,1.5,2,2.5,3)
            spdata$cex=as.character(cut(abs(spdata$resid), 5, labels=cex_))
          plot(spdata$x[spdata$resid<0], spdata$y[spdata$resid<0],col=spdata$col[spdata$resid<0], cex=as.numeric(spdata$cex[spdata$resid<0]), pch= 16, main=list('Australia residuals <0'), xlab = 'Longitude', ylab = 'Latitude')
          plot(spdata$x[spdata$resid>=0], spdata$y[spdata$resid>=0],col=spdata$col[spdata$resid>=0], cex=as.numeric(spdata$cex[spdata$resid>=0]), pch= 16, main=list('Australia residuals >=0'), xlab = 'Longitude', ylab = 'Latitude')
          }
       
       mtext(stringr::str_wrap(paste(paste0(name," model: "), slot(mo,"call")[1],'(',slot(mo,"call")[2],sep=''), width = ceiling(nchar(paste(slot(mo,"call")[1],'(',slot(mo,"call")[2],sep=''))/2)+10), side = 3, line = 0, cex=0.5,outer = TRUE, col = 'darkblue') #ceiling(nchar(paste(slot(mo,"call")[1],'(',slot(mo,"call")[2],sep=''))/2)
       if(PNG==TRUE){dev.off()}
      }
    
  
 # data
    ph  =  fread(here::here('Data/phylopic.txt'))
    setnames(ph, old = c('Name', 'Code'), new = c('genus2', 'uid'))

    t = fread(here::here('Data/taxonomy.txt'))
    
    g = fread(here::here('Data/google_mobility.txt')) #fwrite(d, here::here('Data/data.txt'), sep ='\t')
    g[, Year := as.integer(substring(date, nchar(date)-3, nchar(date)))]
    g[nchar(date)==9, date:=paste0('0',date)]
    g[, date_ :=as.Date(date, format = '%d.%m.%Y')]
    g[, Day :=yday(date_)]
    g[country_region!='Australia', Day := Day-92 +1] # 1 April = start of breeding season (1st day) = 92 day of the year 
    g[country_region=='Australia', Day := Day-228 +1] # 15 Augusst = start of breeding season (1st day) = 228 day of the year 
    setnames(g, old = 'country_region', new ='Country')

    d = fread(here::here('Data/data_corrected.txt')) #fwrite(d, here::here('Data/data.txt'), sep ='\t')
    # add data and weekdays
    x = fread(here::here("Data/date.txt"))[, .(IDObs, Date_corr)]
    x[, date_:=as.Date(Date_corr, "%d.%m.%Y" )]
    x[, weekday := weekdays(date_)]
    d =  merge(d, x[, .(IDObs, date_, weekday)], by = "IDObs")
    #d[, date_ := as.Date(Day, origin = paste0(Year, '-01-01'))]

    # adjust correct assignment of season (Year) for Australia
    d[Country == 'Australia' & Year == 2020 & Covid == 0, Year:=2019]
    d[Country == 'Australia' & Year == 2021 & Day>139, Year:=2020]
    d[Country == 'Australia' & Year == 2022 & Day>139, Year:=2021]

    d = d[order(Year, IDLocality, Day, Hour)]
    d[Country %in% c("Czech_Republic", "Czech Republic"), Country := "Czechia"]
    d[, genus := sub("_.*", "", Species)]
    d[, sp_day_year := paste(Year, Species, Day, sep="_")]
    d[, sp_loc := paste(Species, IDLocality, sep="_")]
    d[, sp_country := paste(Species, Country, sep="\n")]
    d[, rad:=(2*pi*Hour) / 24]
    d[, Day_:= Day]
    #d[, FID_z := scale(FID), by = Species]
    #d[, SD_z := scale(SD), by = Species]
    d[, FID_ln := log(FID)]
    d[, SD_ln := log(SD)]
    d[, body_ln := log(BodyMass)]
    d[, flock_ln := log(FlockSize)]
    d[, weekday := weekdays(date_)]
    #d[Country == 'Australia', Day_:= abs(Day - 189)]
    
    d1 = d[Covid == 1, .N, by = Species]
    d2 = d[Covid == 0, .N, by = Species]
    setnames(d1, old = 'N', new ='N_during')
    setnames(d2, old = 'N', new ='N_before')
    dd = merge(d1,d2) # species with data before and during
    da = merge(d1,d2, all = TRUE)

    d1p = d[Country!='Poland' & Covid == 1, .N, by = Species]
    d2p = d[Country!='Poland' & Covid == 0, .N, by = Species]
    setnames(d1p, old = 'N', new ='N_during')
    setnames(d2p, old = 'N', new ='N_before')
    ddp = merge(d1p,d2p) # species with data before and during, but without Poland data

    p1 = d[Covid == 1, .N, by = .(IDLocality, Species)]
    p2 = d[Covid == 0, .N, by = .(IDLocality, Species)]
    setnames(p1, old = 'N', new ='N_during')
    setnames(p2, old = 'N', new ='N_before')
    pp = merge(p1,p2)  # species-localities with data before and during
    pa = merge(p1,p2, all = TRUE)

    p1p = d[Country!='Poland' & Covid == 1, .N, by = .(IDLocality, Species)]
    p2p = d[Country!='Poland' & Covid == 0, .N, by = .(IDLocality, Species)]
    setnames(p1p, old = 'N', new ='N_during')
    setnames(p2p, old = 'N', new ='N_before')
    ppp = merge(p1p,p2p)  # species-localities with data before and during,but without Poland data 
   
    p1p4 = d[Year!=2014 & Country!='Poland' & Covid == 1, .N, by = .(IDLocality, Species)]
    p2p4 = d[Year!=2014 &Country!='Poland' & Covid == 0, .N, by = .(IDLocality, Species)]
    setnames(p1p4, old = 'N', new ='N_during')
    setnames(p2p4, old = 'N', new ='N_before')
    ppp4 = merge(p1p4,p2p4) # species-localities with data before and during, but without Poland & 2014 data

    s = d[Covid == 1]
    s[, Nsp := .N, by ='Species']
    s[, sp := gsub('[_]', ' ', Species)]
    # add google mobility
    s = merge(s, g[,.(Country,  date_, parks_percent_change_from_baseline)], all.x = TRUE)
    s[, year_ := as.character(Year)]  


    ss = s[!is.na(parks_percent_change_from_baseline)]
    ss[, country_year := paste(Country, Year)] #table(paste(s$Country, s$Year))   
    ss[parks_percent_change_from_baseline<0, google := 'before_zero']
    ss[parks_percent_change_from_baseline>0, google := 'after_zero']
    ss[, sp_country_google:= paste(sp_country, google)]

    g[, weekday := weekdays(date_)]
```

## Title: Urban birds' flight responses were largely unaffected by the COVID-19 shutdowns
## correlations among predictors

```{r cor_pred, fig.width=7.5, fig.height = 7.5}
  d[, sin_rad:=sin(rad)]
  d[, cos_rad := cos(rad)]

  dp = d[, c("SD_ln", "flock_ln", "body_ln", "sin_rad", "cos_rad", "Temp", "Day")]
  setnames(dp, old = c("SD_ln", "flock_ln", "body_ln", "sin_rad", "cos_rad", "Temp", "Day"), new = c("Starting distance\nln(m)", "Flock size\nln(m)", "Body mass\nln(m)", "Sine\nof radians", "Cosine\nof radians", "Temperature\n°C", "Day"))

  png("Outputs/Fig_S1_rev.png", width = 19, height = 19, units = "cm", bg = "transparent", res = 600)
  chart.Correlation(dp, histogram = TRUE, pch = 19, alpha = 0.5)
  mtext("Single observations", side = 3, line = 3)
  dev.off()
```

## prediction for FID ~ Period

```{r }
# predictions
# full
ms <- lmer(scale(log(FID)) ~
    scale(log(SD)) +
    scale(log(FlockSize)) +
    scale(log(BodyMass)) +
    scale(sin(rad)) + scale(cos(rad)) +
    # scale(Day)+
    scale(Temp) +
    scale(Covid) +
    (1 | Year) + (1 | weekday) + (1| genus) + (1| Species) + (1 | sp_day_year) + (scale(Covid) | Country) + (scale(Covid) | IDLocality) + (1 | sp_loc),
data = d, REML = FALSE,
control <- lmerControl(
    optimizer = "optimx", optCtrl = list(method = "nlminb")
)
)
est_ms <- est_out(ms, "All: (1|Year) + (1|genus) + (1|Species) + (1|sp_day_year) + (scale(Covid)|Country) + (scale(Covid)|IDLocality) + (1|sp_loc)")
est_ms[, control_for_starting_distance := "yes"]

mx <- lmer(scale(log(FID)) ~
    #scale(log(SD)) +
    scale(log(FlockSize)) +
    scale(log(BodyMass)) +
    scale(sin(rad)) + scale(cos(rad)) +
    # scale(Day)+
    scale(Temp) +
    scale(Covid) +
    (1 | Year) + (1 | weekday) + (1 | genus) + (1 | Species) + (1 | sp_day_year) + (scale(Covid) | Country) + (scale(Covid) | IDLocality) + (1 | sp_loc),
    data = d, REML = FALSE,
        control = lmerControl( 
            optimizer ='optimx', optCtrl=list(method='nlminb')) 
        )  
# (Covid|IDLocality) +
est_mx <- est_out(mx, "All: (1|Year) + (1|genus) + (1|Species) + (1|sp_day_year) + (scale(Covid)|Country) + (scale(Covid)|IDLocality) + (1|sp_loc)")
est_mx[, control_for_starting_distance := "no"]

# CZ - singular fits only due to genera estimated as zero (removing it changes no results)
  cs <- lmer(scale(log(FID)) ~
    scale(log(SD)) +
    scale(log(FlockSize)) +
    scale(log(BodyMass)) +
    scale(sin(rad)) + scale(cos(rad)) +
    # scale(Day)+
    scale(Temp) +
    scale(Covid) +
    (1 | Year) + (1 | weekday) + (1 | genus) + (1| Species) + (1 | sp_day_year) + (scale(Covid)| IDLocality) + (1|sp_loc),
    #(1 | Year) + (1 | weekday) + (1|genus) + (1 | Species) + (1 | sp_day_year) + (1 | IDLocality),
    data = d[Country == "Czechia"], REML = FALSE
    )
  est_cs <- est_out(cs, "Czechia: (1|weekday)+(1|genus)+(1|Species)+(1|sp_day_year)+(scale(Covid)|IDLocality)+(1|sp_loc)")
  est_cs[, control_for_starting_distance := "yes"]
  
  cx <- lmer(scale(log(FID)) ~
    #scale(log(SD)) +
    scale(log(FlockSize)) +
    scale(log(BodyMass)) +
    scale(sin(rad)) + scale(cos(rad)) +
    # scale(Day)+
    scale(Temp) +
    scale(Covid) +
    (1 | Year) + (1 | weekday) + (1 | genus) + (1| Species) + (1 | sp_day_year) + (scale(Covid)| IDLocality) + (1|sp_loc),
    #(1 | Year) + (1 | weekday) + (1 | genus) + (1 | Species) + (1 | sp_day_year) + (1 | IDLocality),
    data = d[Country == "Czechia"], REML = FALSE
    )
  est_cx <- est_out(cx, "Czechia: (1|weekday)+(1|genus)+(1|Species)+(1|sp_day_year)+(scale(Covid)|IDLocality)+(1|sp_loc)")
  est_cx[, control_for_starting_distance := "no"]

# FI
fs <- lmer(scale(log(FID)) ~
    scale(log(SD)) +
    scale(log(FlockSize)) +
    scale(log(BodyMass)) +
    scale(sin(rad)) + scale(cos(rad)) +
    # scale(Day)+
    scale(Temp) +
    scale(Covid) +
    (1 | Year) + (1 | weekday) + (1 | genus) + (1| Species) + (1 | sp_day_year) + (scale(Covid)| IDLocality) + (1|sp_loc),
data = d[Country == "Finland"], REML = FALSE
)
est_fs <- est_out(fs, "Finland: (1|weekday)+(1|genus)+(1|Species)+(1|sp_day_year)+(scale(Covid)|IDLocality)+(1|sp_loc)")
est_fs[, control_for_starting_distance := "yes"]

fx <- lmer(scale(log(FID)) ~
    # scale(log(SD)) +
    scale(log(FlockSize)) +
    scale(log(BodyMass)) +
    scale(sin(rad)) + scale(cos(rad)) +
    # scale(Day)+
    scale(Temp) +
    scale(Covid) +
     (1 | Year) + (1 | weekday) + (1 | genus) + (1| Species) + (1 | sp_day_year) + (scale(Covid)| IDLocality) + (1|sp_loc),
data = d[Country == "Finland"], REML = FALSE
)
est_fx <- est_out(fx, "Finland: (1|weekday)+(1|genus)+(1|Species)+(1|sp_day_year)+(scale(Covid)|IDLocality)+(1|sp_loc)")
est_fx[, control_for_starting_distance := "no"]

# HU - singular fits only due to sp_loc estimated as zero (removing it changes no results)
hs <- lmer(scale(log(FID)) ~
    scale(log(SD)) +
    scale(log(FlockSize)) +
    scale(log(BodyMass)) +
    scale(sin(rad)) + scale(cos(rad)) +
    # scale(Day)+
    scale(Temp) +
    scale(Covid) +
    (1 | Year) + (1 | weekday) + (1 | genus) + (1 | Species) + (1 | sp_day_year) + (scale(Covid) | IDLocality) + (1 | sp_loc),
# (1 | Year) + (1 | weekday) + (1|genus) + (1 | Species) + (1 | sp_day_year) + (1 | IDLocality),
data = d[Country == "Hungary"], REML = FALSE
)
est_hs <- est_out(hs, "Hungary: (1|weekday)+(1|genus)+(1|Species)+(1|sp_day_year)+(scale(Covid)|IDLocality)+(1|sp_loc)")
est_hs[, control_for_starting_distance := "yes"]

hx <- lmer(scale(log(FID)) ~
    # scale(log(SD)) +
    scale(log(FlockSize)) +
    scale(log(BodyMass)) +
    scale(sin(rad)) + scale(cos(rad)) +
    # scale(Day)+
    scale(Temp) +
    scale(Covid) +
    (1 | Year) + (1 | weekday) + (1 | genus) + (1 | Species) + (1 | sp_day_year) + (scale(Covid) | IDLocality) + (1 | sp_loc),
# (1 | Year) + (1 | weekday) + (1 | genus) + (1 | Species) + (1 | sp_day_year) + (1 | IDLocality),
data = d[Country == "Hungary"], REML = FALSE
)
est_hx <- est_out(hx, "Hungary: (1|weekday)+(1|genus)+(1|Species)+(1|sp_day_year)+(scale(Covid)|IDLocality)+(1|sp_loc)")
est_hx[, control_for_starting_distance := "no"]

# AU - singular fits only due to Year and random slope estimated as zero (removing those changes no results)
as <- lmer(scale(log(FID)) ~
    scale(log(SD)) +
    scale(log(FlockSize)) +
    scale(log(BodyMass)) +
    scale(sin(rad)) + scale(cos(rad)) +
    # scale(Day)+
    scale(Temp) +
    scale(Covid) +
     (1 | Year) +(1 | weekday) + (1 | genus) + (1 | Species) + (1 | sp_day_year) + (scale(Covid) | IDLocality) + (1 | sp_loc),
# (1 | Year) + (1 | weekday) + (1|genus) + (1 | Species) + (1 | sp_day_year) + (1 | IDLocality),
data = d[Country == "Australia"], REML = FALSE
)
est_as <- est_out(as, "Australia: (1|weekday)+(1|genus)+(1|Species)+(1|sp_day_year)+(scale(Covid)|IDLocality)+(1|sp_loc)")
est_as[, control_for_starting_distance := "yes"]

ax <- lmer(scale(log(FID)) ~
    # scale(log(SD)) +
    scale(log(FlockSize)) +
    scale(log(BodyMass)) +
    scale(sin(rad)) + scale(cos(rad)) +
    # scale(Day)+
    scale(Temp) +
    scale(Covid) +
    (1 | Year) + (1 | weekday) + (1 | genus) + (1 | Species) + (1 | sp_day_year) + (scale(Covid) | IDLocality) + (1 | sp_loc),
# (1 | Year) + (1 | weekday) + (1 | genus) + (1 | Species) + (1 | sp_day_year) + (1 | IDLocality),
data = d[Country == "Australia"], , REML = FALSE,
        control = lmerControl( 
            optimizer ='optimx', optCtrl=list(method='nlminb')) 
)
est_ax <- est_out(ax, "Australia: (1|weekday)+(1|genus)+(1|Species)+(1|sp_day_year)+(scale(Covid)|IDLocality)+(1|sp_loc)")
est_ax[, control_for_starting_distance := "no"]

# PL 
ps <- lmer(scale(log(FID)) ~
    scale(log(SD)) +
    scale(log(FlockSize)) +
    scale(log(BodyMass)) +
    scale(sin(rad)) + scale(cos(rad)) +
    # scale(Day)+
    scale(Temp) +
    scale(Covid) +
    (1 | Year) + (1 | weekday) + (1 | genus) + (1 | Species) + (1 | sp_day_year),
# (1 | Year) + (1 | weekday) + (1|genus) + (1 | Species) + (1 | sp_day_year) + (1 | IDLocality),
data = d[Country == "Poland"], REML = FALSE
)
est_ps <- est_out(ps, "Poland: (1|weekday)+(1|genus)+(1|Species)+(1|sp_day_year)")
est_ps[, control_for_starting_distance := "yes"]

px <- lmer(scale(log(FID)) ~
    # scale(log(SD)) +
    scale(log(FlockSize)) +
    scale(log(BodyMass)) +
    scale(sin(rad)) + scale(cos(rad)) +
    # scale(Day)+
    scale(Temp) +
    scale(Covid) +
    (1 | Year) + (1 | weekday) + (1 | genus) + (1 | Species) + (1 | sp_day_year),
# (1 | Year) + (1 | weekday) + (1 | genus) + (1 | Species) + (1 | sp_day_year) + (1 | IDLocality),
data = d[Country == "Poland"], REML = FALSE
)
est_px <- est_out(px, "Poland: (1|weekday)+(1|genus)+(1|Species)+(1|sp_day_year)")
est_px[, control_for_starting_distance := "no"]

  # combine
    est_ms[, Country := 'All\n(mixed model)']
    est_mx[, Country := "All\n(mixed model)"]
    est_as[, Country := "Australia"]
    est_ax[, Country := "Australia"]
    est_cs[, Country := "Czechia"]
    est_cx[, Country := "Czechia"]
    est_hs[, Country := "Hungary"]
    est_hx[, Country := "Hungary"]
    est_ps[, Country := "Poland"]
    est_px[, Country := "Poland"]
    est_fs[, Country := "Finland"]
    est_fx[, Country := "Finland"]

    o = rbind(est_ms, est_mx, 
            est_as, est_ax, 
            est_cs, est_cx, 
            est_hs, est_hx,
            est_ps, est_px, 
            est_fs, est_fx)
    save(o, file = here::here('Data/dat_est_rev.Rdata'))
 
  #AIC(hs, hx)
  #AIC(cs, cx)
  #0AIC(ps, px)
  #AIC(as, ax)
  #AIC(fs, fx)

```
```{r est_1, fig.width=3, fig.height = 2.5}
load(here::here("Data/dat_est_rev.Rdata"))
o[predictor %in% c("scale(Covid)"), predictor := "Period"]
oo <- o[predictor %in% c("Period")]
oo[, N:=as.numeric(sub('.*N = ', '', model))]
# add meta-analytical mean
  oo_s = oo[control_for_starting_distance == 'yes']
  met = summary(meta.summaries(d = oo_s$estimate, se = oo_s$sd, method = "fixed", weights = oo_s$N))$summci
  oo_met = data.table(predictor = "Period", estimate = met[2], lwr = met[1], upr = met[3], sd = NA, model = NA, control_for_starting_distance = "yes", Country = "Combined\n(metanalytical)", N = NA)

  oo_sx = oo[control_for_starting_distance == "no"]
  metx = summary(meta.summaries(d = oo_sx$estimate, se = oo_sx$sd, method = "fixed", weights = oo_sx$N))$summci
  oo_metx = data.table(predictor = "Period", estimate = metx[2], lwr = metx[1], upr = metx[3], sd = NA, model = NA, control_for_starting_distance = "no", Country = "Combined\n(metanalytical)", N = NA)
  
  oo = rbind(oo, oo_met, oo_metx)
    
oo[, Country := factor(Country, levels = rev(c("Finland", "Poland", "Czechia", "Hungary", "Australia", "Combined\n(metanalytical)", "All\n(mixed model)")))]

width_ <- .5 # spacing between error bars

#col_ <- c(brewer.pal(n = 12, name = "Paired"), "grey30", "grey80")
#Tol_bright <- c("#EE6677", "#228833", "#4477AA", "#CCBB44", "#66CCEE", "#AA3377", "#BBBBBB")
#Tol_muted <- c("#88CCEE", "#44AA99", "#117733", "#332288", "#DDCC77", "#999933", "#CC6677", "#882255", "#AA4499", "#DDDDDD")
#Tol_light <- c("#BBCC33", "#AAAA00", "#77AADD", "#EE8866", "#EEDD88", "#FFAABB", "#99DDFF", "#44BB99", "#DDDDDD")

# From Color Universal Design (CUD): https://jfly.uni-koeln.de/color/
#Okabe_Ito <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#000000")
#col_ = Okabe_Ito[7:1]
# JAMA and LocusZoom modified order
#col_ =  c("#374E55FF", "#374E55FF", "#DF8F44FF", "#79AF97FF", "#00A1D5FF", "#B24745FF",  "#80796BFF") #"#6A6599FF",
#col_ <- c("#357EBDFF", "#9632B8FF", "#46B8DAFF", "#5CB85CFF", "#EEA236FF", "#D43F3AFF", "#D43F3AFF")[7:1] # "#D43F3AFF", "#B8B8B8FF"
col_ = c("#357EBDFF", "#D43F3AFF", "#46B8DAFF", "#5CB85CFF", "#EEA236FF", "#9632B8FF", "#9632B8FF")[7:1] # "#D43F3AFF", "#B8B8B8FF"
#show_col(col_)

#p = 
ggplot(oo, aes(x = estimate, y = Country, col = Country, shape = control_for_starting_distance)) +
    geom_vline(xintercept = 0, color = "grey", linetype = "dotted") +
    geom_errorbarh(aes(xmin = lwr, xmax = upr), height = 0, position = ggstance::position_dodgev(width_)) +
    # geom_point(position = ggstance::position_dodgev(.6)) +
    geom_point(position = position_dodge(width = width_), bg = "white", size = 1.1) +
    # scale_color_viridis(discrete=TRUE, begin=0, end = 0.5)  +
    # scale_fill_viridis(discrete=TRUE, begin=0, end = 0.5) +
    # geom_text( aes(x = n_pos,label = N), vjust = 0, size = 1.75, position = ggstance::position_dodgev(width_))+ # 3 positions for 3 bars
    # annotate("text", x=log10(3), y=85, label= "Used", col = "grey30", size = 2.5)+

    scale_shape_manual(name = "Controlled for\nstarting distance", guide = guide_legend(reverse = TRUE), values = c(21, 19)) +
    #scale_color_jama(guide = "none")+ #, palette = 'light'
    scale_color_manual(guide = "none", values = col_) + #guide_legend(reverse = TRUE)
    scale_x_continuous(breaks = round(seq(-0.6, 1.2, by = 0.3), 1)) +
    ylab("") +
    xlab("Standardized effect size of period\n[on flight initiation distance]") +
    # coord_cartesian(xlim = c(-.15, .15)) +
    # scale_x_continuous(breaks = round(seq(-.15, .15, by = 0.05),2)) +
    theme_bw() +
    theme(
        legend.position = "right",
        legend.title = element_text(size = 7),
        legend.text = element_text(size = 6),
        # legend.spacing.y = unit(0.1, 'cm'),
        legend.key.height = unit(0.5, "line"),
        legend.margin = margin(0, 0, 0, 0),
        # legend.position=c(0.5,1.6),
        plot.title = element_text(color = "grey", size = 7),
        plot.margin = margin(b = 0.5, l = 0.5, t = 0.5, r = 0.5, unit = "pt"),
        panel.grid = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = ax_lines, size = 0.25),
        axis.line.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.ticks.x = element_line(colour = ax_lines, size = 0.25),
        # axis.text.x = element_text()
        axis.ticks.length = unit(1, "pt"),
        axis.text.x = element_text(, size = 6),
        axis.text.y = element_text(colour = "black", size = 7),
        axis.title = element_text(size = 7)
    )

ggsave(here::here("Outputs/Fig_rev_width_CustomLocusZoom_v2.png"), width = 8, height = 6, unit = "cm", dpi = 600)
#ggsave("Outputs/Fig_rev_width_CustomJAMAv1.png", width = 8, height = 6, unit = "cm", dpi = 600)
#ggsave("Outputs/Fig_rev_width_Okabe_v2.png", width = 8, height = 6, unit = "cm", dpi = 600)
#ggsave("Outputs/Fig_rev_width_UChicago_v3.png", width = 8, height = 6, unit = "cm", dpi = 600)
#ggsave("Outputs/Fig_rev_width_d3_v2.png", width = 8, height = 6, unit = "cm", dpi = 600)
#ggsave("Outputs/Fig_rev_width_nejm_v2.png", width = 8, height = 6, unit = "cm", dpi = 600)
#ggsave("Outputs/Fig_rev_width_jama_v2.png", width = 8, height = 6, unit = "cm", dpi = 600)
#ggsave("Outputs/Fig_rev_width_jco_v2.png", width = 8, height = 6, unit = "cm", dpi = 600)
#ggsave("Outputs/Fig_rev_width_npg_v2.png", width = 8, height = 6, unit = "cm", dpi = 600)

# show used colors
#gg <- ggplot_build(p)
#col_ = unique(gg$data[[3]]["colour"])$colour
#show_col(col_)
```

**Fig. 1 | The effect size with 95%CIs of Period (0 – before, 1 – during shutdowns) on flight initiation distance (ln-transformed).**  The general model on all data was controlled for starting distance (ln-transformed; filled circles) or not (empty circles), flock size (ln-transformed), temperature (also a proxy for a day within the breeding season: r Pearson = 0.48; Fig. S1) and time of day. To account for circular properties of time, time was transformed into radians (2 × time × π/24) and fitted as sine and cosine of radians (Bulla et al. 2016). All continuous variables were standardised by subtracting the mean and dividing by the standard deviation. The multicollinearity was small as correlations between predictors were  weak (r<XX, Fig. S1). To account for the non-independence of data points (Schielzeth & Forstmeier 2009; Barr et al. 2013), we attempted to fit random intercepts of year, weekday, genus, species, species at a given day and year, country, site, and species within a site, while fitting Period as random slope within site. Note that fitting Period as random slope at other random intercepts produces similar results (see Fig Sxx). We used this approach with a full dataset with all observations (n = 6369), as well as with conservative datasets, one with at least five observations per species and Period (i.e. at least five observations before and five during the COVID-19 shutdowns; n = 5260), the other with at least 10 observations per species and each Period (n = 5106). In other words, conservative datasets included species sampled in the both Periods. Albeit random structure of the full model accounts for the potential country specific effect (country accounted for xx% of variance, Table xx), depicted are also estimates from country specific models, with same predictors and random structure as the full model (excluding the random intercept of country and in case of Poland also site, as specific sites were not noted in Poland). Using meta.summaries function from rmeta R-package we used the country estimates, their standard deviation, and sample size per country to estimate the meta-analytical mean, which reflects the results based on the full mixed effect model. Also, albeit the response of humans to shutdowns were likely country specific and presence of humans in parks might have increased in some countries, decreased in others or did not change, the escape distances do not reflect such changes. In other words, the birds are either inflexible or the change in human behavior due to shutdowns was not strong enought, which might have been the case - see Fig. ZZ - decide what to show in this figure (I think 2020-2022 changes in Google Mobility for each country including the fits and refering to supplementary figure)


```{r line, fig.width=10, fig.height = 6}
g_ <- fread(here::here("Data/google_mobility.txt")) # fwrite(d, here::here('Data/data.txt'), sep ='\t')
g_[, Year := as.integer(substring(date, nchar(date) - 3, nchar(date)))]
g_[nchar(date) == 9, date := paste0("0", date)]
g_[, date_ := as.Date(date, format = "%d.%m.%Y")]
g_[, Day := yday(date_)]
setnames(g_, old = "country_region", new = "Country")
g_[, weekday := weekdays(date_)]

g0 = ggplot(g_, aes(x = parks_percent_change_from_baseline, fill = factor(Year))) +
  geom_histogram(position = "dodge") +
  # scale_y_continuous(trans = 'log')+
  scale_fill_manual(values = c("orange", "skyblue", "black"), guide = 'none') +
  geom_vline(xintercept = 0, lty = 3, col = "red") +
  labs(subtitle = "Distribution") +
  xlab( 'Google Mobility\n[% change in human presence]') +
  ylab( 'Count') +
  facet_wrap(~Country, nrow = 5)

g1 = ggplot(g_, aes(x = Day, y = parks_percent_change_from_baseline, col = factor(Year))) +
  geom_line() +
  facet_wrap(~Country, nrow = 5) +
  labs(subtitle = "Raw data", xlab = 'Day\n ', y = 'Google Mobility\n[% change in human presence]') +
  # scale_y_continuous(trans = 'log')+
  coord_cartesian(ylim = c(-100, 300))+
  scale_color_manual(values = c("orange", "skyblue", "black"), guide = 'none')

g2 = ggplot(g_, aes(x = Day, y = parks_percent_change_from_baseline, col = factor(Year))) +
     stat_smooth() +
     facet_wrap(~Country, nrow = 5) +
     labs(subtitle = "Loess", xlab = 'Day\n ') +
     # scale_y_continuous(trans = 'log')+
     coord_cartesian(ylim = c(-100, 300))+
     scale_color_manual(values = c("orange", "skyblue", "black"), name = 'Year')+
      theme(
         axis.text.y = element_blank(),
         axis.title.y = element_blank()
       )
  ggarrange(
    g0, g1, g2,
    ncol = 3, widths = c(0.9, 1, 1.1)
  )
 ggsave(here::here("Outputs/Fig_ZZ.png"), width = 8*2.54, height = 5*2.54, unit = "cm", dpi = 600)
```

**Fig. ZZ | Changes in human presence (Google Mobility) in parks across year and between years.** Left plots represent the raw data, right plots LOESS smoothed curves. Google Mobility is absent for years before COVID-19. Nevertheless, 2022 was a year without shutdowns in the studied countries. Assuming that human activities levels might have been similar to pre-COVID-19 years (which may not be the case), human acctivity in the shutdown years (2020 and 2021) decreased. However, such decrease might be irrelevant for birds as the day-to-day variatioin in human presence seems larger than the general decrease in activity. For weekday specific plot see Fig. S_ZZ


```{r gm_week_year, fig.width=8, fig.height = 6}
g[, weekday := factor(weekday, levels = (c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")))]

ggplot(g, aes(x = Day, y = parks_percent_change_from_baseline, col = factor(Year))) +
    geom_line() +
    facet_grid(rows = vars(Country), cols = vars(weekday)) +
    # scale_y_continuous(trans = 'log')+
    scale_color_manual(values = c("orange", "skyblue", "black"))
```

**Fig. S_ZZ | Changes in human presence (Google Mobility) in parks across weekdays and years.** Depicted are  raw data.
##  Google Mobility vs Stringency

```{r gsfig_mod, fig.width=3, fig.height = 2.5}
# Predictions 
l = list()
ll = list()
 sc = s[Country == "Czechia"]
 cz <- lmer(parks_percent_change_from_baseline ~
    StringencyIndex + 
   (scale(StringencyIndex)|weekday),
 # (1 | Year) + (1 | weekday) + (1|genus) + (1 | Species) + (1 | sp_day_year) + (1 | IDLocality),
 data = sc, REML = FALSE
 )

 ll[[3]] = m_out(name = "Table S_g-s - CZ", dep = "Google Mobility", model = cz, nsim = 5000)

 bsim <- sim(cz, n.sim = nsim)
 v <- apply(bsim@fixef, 2, quantile, prob = c(0.5))
 newD <- data.frame(StringencyIndex = seq(min(sc$StringencyIndex), max(sc$StringencyIndex), length.out = 100)) # values to predict for
 X <- model.matrix(~StringencyIndex, data = newD) # exactly the model which was used has to be specified here
 newD$pred <- (X %*% v)
 predmatrix <- matrix(nrow = nrow(newD), ncol = nsim)
 for (j in 1:nsim) predmatrix[, j] <- (X %*% bsim@fixef[j, ])
 newD$lwr <- apply(predmatrix, 1, quantile, prob = 0.025)
 newD$upr <- apply(predmatrix, 1, quantile, prob = 0.975)
 newD$pred <- apply(predmatrix, 1, quantile, prob = 0.5)
 newD$Country = 'Czechia'
 l[[1]] = newD
 
 s[, year_weekday :=paste(Year, weekday)]
 sf = s[Country == "Finland"]
 fi <- lmer(parks_percent_change_from_baseline ~
    Year+
    StringencyIndex + 
   (scale(StringencyIndex)|year_weekday),
 # (1 | Year) + (1 | weekday) + (1|genus) + (1 | Species) + (1 | sp_day_year) + (1 | IDLocality),
 data = sf, REML = FALSE
 )
 ll[[1]] = m_out(name = "Table S_g-s - FI", dep = "Google Mobility", model = fi, nsim = 5000)

 bsim <- sim(fi, n.sim = nsim)
 v <- apply(bsim@fixef, 2, quantile, prob = c(0.5))
 newD <- data.frame(Year = mean(sf$Year), StringencyIndex = seq(min(sf$StringencyIndex), max(sf$StringencyIndex), length.out = 100)) # values to predict for
 X <- model.matrix(~Year + StringencyIndex, data = newD) # exactly the model which was used has to be specified here
 newD$pred <- (X %*% v)
 predmatrix <- matrix(nrow = nrow(newD), ncol = nsim)
 for (j in 1:nsim) predmatrix[, j] <- (X %*% bsim@fixef[j, ])
 newD$lwr <- apply(predmatrix, 1, quantile, prob = 0.025)
 newD$upr <- apply(predmatrix, 1, quantile, prob = 0.975)
 newD$pred <- apply(predmatrix, 1, quantile, prob = 0.5)
 newD$Country = 'Finland'
 newD$Year = NULL
 l[[2]] = newD

 sh <- s[Country == "Hungary"]
 hu <- lmer(parks_percent_change_from_baseline ~
     Year +
     StringencyIndex +
     (scale(StringencyIndex) | year_weekday),
 # (1 | Year) + (1 | weekday) + (1|genus) + (1 | Species) + (1 | sp_day_year) + (1 | IDLocality),
 data = sh, REML = FALSE
 )
ll[[4]] = m_out(name = "Table S_g-s - HU", dep = "Google Mobility", model = hu, nsim = 5000)
 bsim <- sim(hu, n.sim = nsim)
 v <- apply(bsim@fixef, 2, quantile, prob = c(0.5))
 newD <- data.frame(Year = mean(sh$Year), StringencyIndex = seq(min(sh$StringencyIndex), max(sh$StringencyIndex), length.out = 100)) # values to predict for
 X <- model.matrix(~ Year + StringencyIndex, data = newD) # exactly the model which was used has to be specified here
 newD$pred <- (X %*% v)
 predmatrix <- matrix(nrow = nrow(newD), ncol = nsim)
 for (j in 1:nsim) predmatrix[, j] <- (X %*% bsim@fixef[j, ])
 newD$lwr <- apply(predmatrix, 1, quantile, prob = 0.025)
 newD$upr <- apply(predmatrix, 1, quantile, prob = 0.975)
 newD$pred <- apply(predmatrix, 1, quantile, prob = 0.5)
 newD$Country <- "Hungary"
 newD$Year <- NULL
 l[[3]] <- newD
 
 sp <- s[Country == "Poland"]
 pl <- lmer(parks_percent_change_from_baseline ~
     Year +
     StringencyIndex +
     (scale(StringencyIndex) | year_weekday),
 # (1 | Year) + (1 | weekday) + (1|genus) + (1 | Species) + (1 | sp_day_year) + (1 | IDLocality),
 data = sp, REML = FALSE
 )
 ll[[2]] = m_out(name = "Table S_g-s - PL", dep = "Google Mobility", model = pl, nsim = 5000)
 bsim <- sim(pl, n.sim = nsim)
 v <- apply(bsim@fixef, 2, quantile, prob = c(0.5))
 newD <- data.frame(Year = mean(sp$Year), StringencyIndex = seq(min(sp$StringencyIndex), max(sp$StringencyIndex), length.out = 100)) # values to predict for
 X <- model.matrix(~ Year + StringencyIndex, data = newD) # exactly the model which was used has to be specified here
 newD$pred <- (X %*% v)
 predmatrix <- matrix(nrow = nrow(newD), ncol = nsim)
 for (j in 1:nsim) predmatrix[, j] <- (X %*% bsim@fixef[j, ])
 newD$lwr <- apply(predmatrix, 1, quantile, prob = 0.025)
 newD$upr <- apply(predmatrix, 1, quantile, prob = 0.975)
 newD$pred <- apply(predmatrix, 1, quantile, prob = 0.5)
 newD$Country <- "Poland"
 newD$Year <- NULL
 l[[4]] <- newD

 sa <- s[Country == "Australia"]
 au <- lmer(parks_percent_change_from_baseline ~
     Year +
     StringencyIndex +
     (scale(StringencyIndex) | year_weekday),
 # (1 | Year) + (1 | weekday) + (1|genus) + (1 | Species) + (1 | sp_day_year) + (1 | IDLocality),
 data = sa, REML = FALSE
 )
 ll[[5]] = m_out(name = "Table S_g-s - AU", dep = "Google Mobility", model = au, nsim = 5000)
 bsim <- sim(au, n.sim = nsim)
 v <- apply(bsim@fixef, 2, quantile, prob = c(0.5))
 newD <- data.frame(Year = mean(sa$Year), StringencyIndex = seq(min(sa$StringencyIndex), max(sa$StringencyIndex), length.out = 100)) # values to predict for
 X <- model.matrix(~ Year + StringencyIndex, data = newD) # exactly the model which was used has to be specified here
 newD$pred <- (X %*% v)
 predmatrix <- matrix(nrow = nrow(newD), ncol = nsim)
 for (j in 1:nsim) predmatrix[, j] <- (X %*% bsim@fixef[j, ])
 newD$lwr <- apply(predmatrix, 1, quantile, prob = 0.025)
 newD$upr <- apply(predmatrix, 1, quantile, prob = 0.975)
 newD$pred <- apply(predmatrix, 1, quantile, prob = 0.5)
 newD$Country <- "Australia"
 newD$Year <- NULL
 l[[5]] <- newD

out_g_s = data.table(do.call(rbind, ll))

sp = data.table(do.call(rbind,l))
sp[, Country := factor(Country, levels = rev(c("Finland", "Poland", "Czechia", "Hungary", "Australia")))]

```
```{r gsfig_pred, fig.width=4, fig.height = 3.5}
col3_ = c("#357EBDFF", "#D43F3AFF", "#46B8DAFF", "#5CB85CFF", "#EEA236FF", "#9632B8FF", "#9632B8FF")[7:1]
col3__ = col3_[3:7]
#p = 
ggplot(sp, aes(x = StringencyIndex, y = pred, col = Country)) +
  geom_ribbon(aes(ymin = lwr, ymax = upr, fill = Country, color = NULL), alpha = .15) +
  geom_jitter(aes(y = parks_percent_change_from_baseline, fill = Country), data = s, pch = 21, col = 'grey20', width = 0.7, height = 3, alpha = 0.5) +
  geom_line(lwd = 1) +
  labs(subtitle = "Mixed model per country predicitons", y = "Google Mobiligy\n[% change from baseline]", x = "Stringency Index") +
   # scale_color_locuszoom()+
   # scale_fill_locuszoom(guide = "none")
  scale_x_continuous(breaks = round(seq(25, 75, by = 25), 1)) +
  scale_y_continuous(breaks = round(seq(-100, 200, by = 50), 1)) +
  #scale_y_continuous(breaks = round(seq(-100, 175, by = 25), 1)) +
  scale_colour_manual(values = col3__, guide = guide_legend(reverse = TRUE, override.aes = list(size = 0)), 
            labels = paste("<span style='color:",
                                   col3__,
                                   "'>",
                                   levels(sp$Country),
                                   "</span>")
            ) +
  scale_fill_manual(values = col3__, guide = "none") +
  theme_bw() +
  theme(
    legend.text = element_markdown(size = 6),
    #legend.position = "right",
    legend.title = element_blank(),
    # legend.spacing.y = unit(0.1, 'cm'),
    legend.key.height = unit(0.5, "line"),
    legend.key.size = unit(0, "line"),
    legend.margin = margin(0, 0, 0, 0),
    legend.box.margin=margin(-10,1,-10,-10),
    # legend.position=c(0.5,1.6),
    plot.title = element_text(color = "grey", size = 7),
    plot.subtitle = element_text(color = "grey60", size = 6),
    plot.margin = margin(b = 0.5, l = 0.5, t = 0.5, r = 0.5, unit = "pt"),
    panel.grid = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(colour = ax_lines, size = 0.25),
    axis.ticks = element_line(colour = ax_lines, size = 0.25),
    # axis.text.x = element_text()
    axis.ticks.length = unit(1, "pt"),
    axis.text = element_text(, size = 6),
    axis.title = element_text(size = 7)
  )

ggsave(here::here("Outputs/Fig_G-S_rev_width_CustomLocusZoom_year-weekday.png"), width = 7, height = 6, unit = "cm", dpi = 600)
```

**Fig. 2 | Relationship betweeen Google Mobility and Stringency Index.** Lines with shaded areas represent predicted relationships from country-specific mixed effect models controlled for the year and non-independence of data points by including weekday within the year as random intercept and Stringency Index as a random slope. Dots represent raw data, jittered to increase visibility. Colors indicate country. The predictions reveal generally negative and week relationship between Gooble Mobility (human activity) in parks and seveareness of governmental restrictions. 

## predictions for FID ~ Stringency

```{r }
# predictions
 # full
 mss <- lmer(scale(log(FID)) ~
     scale(Year) +
     scale(log(SD)) +
     scale(log(FlockSize)) +
     scale(log(BodyMass)) +
     scale(sin(rad)) + scale(cos(rad)) +
     # scale(Day)+
     scale(Temp) +
     scale(StringencyIndex) +
     (1 | genus) + (1 | Species) + (1 | sp_day_year) + (scale(StringencyIndex) | Country) + (1 | IDLocality) + (1 | sp_loc),
 data = s, REML = FALSE,
 control = lmerControl(
     optimizer = "optimx", optCtrl = list(method = "nlminb")
 )
 )
 est_mss <- est_out(mss, "ALL: (1|genus) + (1|Species) + (1|sp_day_year) + (scale(StringencyIndex)|Country) + (1|IDLocality) +(1|sp_loc)")
 est_mss[, control_for_starting_distance := "yes"]

 msx <- lmer(scale(log(FID)) ~
    scale(Year) +
    #scale(log(SD)) +
    scale(log(FlockSize)) +
    scale(log(BodyMass)) +
    scale(sin(rad)) + scale(cos(rad)) +
    # scale(Day)+
    scale(Temp) +
    scale(StringencyIndex) +
    (1 | genus) + (1 | Species) + (1 | sp_day_year) + (scale(StringencyIndex) | Country) + (1| IDLocality) + (1 | sp_loc),
 data = s, REML = FALSE,
 control = lmerControl(
    optimizer = "optimx", optCtrl = list(method = "nlminb")
)
)
est_msx <- est_out(msx, "ALL: (1|genus) + (1|Species) + (1|sp_day_year) + (scale(StringencyIndex)|Country) + (1|IDLocality) +(1|sp_loc)")
est_msx[, control_for_starting_distance := "no"]


# CZ - singular fits only due to genera estimated as zero (removing it changes no results)
  css <- lmer(scale(log(FID)) ~
    #scale(Year) +
    scale(log(SD)) +
    scale(log(FlockSize)) +
    scale(log(BodyMass)) +
    scale(sin(rad)) + scale(cos(rad)) +
    # scale(Day)+
    scale(Temp) +
    scale(StringencyIndex) +
    (1 | genus) + (1| Species) + (1 | sp_day_year) + (1| IDLocality) + (1|sp_loc),
    #(1 | Year) + (1 | weekday) + (1|genus) + (1 | Species) + (1 | sp_day_year) + (1 | IDLocality),
    data = s[Country == "Czechia"], REML = FALSE
    )
  est_css <- est_out(css, "Czechia: (1|genus)+(1|Species)+(1|sp_day_year)+(1|IDLocality)+(1|sp_loc)")
  est_css[, control_for_starting_distance := "yes"]
  
  csx <- lmer(scale(log(FID)) ~
    #scale(Year) +
    scale(log(SD)) +
    scale(log(FlockSize)) +
    scale(log(BodyMass)) +
    scale(sin(rad)) + scale(cos(rad)) +
    # scale(Day)+
    scale(Temp) +
    scale(StringencyIndex) +
    (1 | genus) + (1| Species) + (1 | sp_day_year) + (1| IDLocality) + (1|sp_loc),
    #(1 | Year) + (1 | weekday) + (1|genus) + (1 | Species) + (1 | sp_day_year) + (1 | IDLocality),
    data = s[Country == "Czechia"], REML = FALSE
    )
  est_csx <- est_out(csx, "Czechia: (1|genus)+(1|Species)+(1|sp_day_year)+(1|IDLocality)+(1|sp_loc)")
  est_csx[, control_for_starting_distance := "no"]

# FI
fss <- lmer(scale(log(FID)) ~
    scale(Year) +
    scale(log(SD)) +
    scale(log(FlockSize)) +
    scale(log(BodyMass)) +
    scale(sin(rad)) + scale(cos(rad)) +
    # scale(Day)+
    scale(Temp) +
    scale(StringencyIndex) +
    (1 | genus) + (1| Species) + (1 | sp_day_year) + (1| IDLocality) + (1|sp_loc),
data = s[Country == "Finland"], REML = FALSE
)
est_fss <- est_out(fss, "Finland: (1|genus)+(1|Species)+(1|sp_day_year)+(scale(StringencyIndex)|IDLocality)+(1|sp_loc)")
est_fss[, control_for_starting_distance := "yes"]

fsx <- lmer(scale(log(FID)) ~
    scale(Year) +
    # scale(log(SD)) +
    scale(log(FlockSize)) +
    scale(log(BodyMass)) +
    scale(sin(rad)) + scale(cos(rad)) +
    # scale(Day)+
    scale(Temp) +
    scale(StringencyIndex) +
    (1 | genus) + (1| Species) + (1 | sp_day_year) + (1| IDLocality) + (1|sp_loc),
data = s[Country == "Finland"], REML = FALSE
)
est_fsx <- est_out(fsx, "Finland: (1|genus)+(1|Species)+(1|sp_day_year)+(scale(StringencyIndex)|IDLocality)+(1|sp_loc)")
est_fsx[, control_for_starting_distance := "no"]

# HU - singular fits only due to sp_loc estimated as zero (removing it changes no results)
hss <- lmer(scale(log(FID)) ~
    scale(Year) +
    scale(log(SD)) +
    scale(log(FlockSize)) +
    scale(log(BodyMass)) +
    scale(sin(rad)) + scale(cos(rad)) +
    # scale(Day)+
    scale(Temp) +
    scale(StringencyIndex) +
     (1 | genus) + (1 | Species) + (1 | sp_day_year) + (1 | IDLocality) + (1 | sp_loc),
# (1 | Year) + (1 | weekday) + (1|genus) + (1 | Species) + (1 | sp_day_year) + (1 | IDLocality),
data = s[Country == "Hungary"], REML = FALSE
)
est_hss <- est_out(hss, "Hungary: (1|genus)+(1|Species)+(1|sp_day_year)+(1|IDLocality)+(1|sp_loc)")
est_hss[, control_for_starting_distance := "yes"]

hsx <- lmer(scale(log(FID)) ~
    scale(Year) +
    # scale(log(SD)) +
    scale(log(FlockSize)) +
    scale(log(BodyMass)) +
    scale(sin(rad)) + scale(cos(rad)) +
    # scale(Day)+
    scale(Temp) +
    scale(StringencyIndex) +
    (1 | genus) + (1 | Species) + (1 | sp_day_year) + (1 | IDLocality) + (1 | sp_loc),
# (1 | Year) + (1 | weekday) + (1 | genus) + (1 | Species) + (1 | sp_day_year) + (1 | IDLocality),
data = s[Country == "Hungary"], REML = FALSE
)
est_hsx <- est_out(hsx, "Hungary: (1|genus)+(1|Species)+(1|sp_day_year)+(1|IDLocality)+(1|sp_loc)")
est_hsx[, control_for_starting_distance := "no"]

# AU - singular fits only due to Year and random slope estimated as zero (removing those changes no results)
ass <- lmer(scale(log(FID)) ~
    scale(Year) +
    scale(log(SD)) +
    scale(log(FlockSize)) +
    scale(log(BodyMass)) +
    scale(sin(rad)) + scale(cos(rad)) +
    # scale(Day)+
    scale(Temp) +
    scale(StringencyIndex) +
      (1 | genus) + (1 | Species) + (1 | sp_day_year) + (1 | IDLocality) + (1 | sp_loc),
# (1 | Year) + (1 | weekday) + (1|genus) + (1 | Species) + (1 | sp_day_year) + (1 | IDLocality),
data = s[Country == "Australia"], REML = FALSE
)
est_ass <- est_out(ass, "Australia: (1|genus)+(1|Species)+(1|sp_day_year)+(1|IDLocality)+(1|sp_loc)")
est_ass[, control_for_starting_distance := "yes"]

asx <- lmer(scale(log(FID)) ~
    scale(Year) +
    # scale(log(SD)) +
    scale(log(FlockSize)) +
    scale(log(BodyMass)) +
    scale(sin(rad)) + scale(cos(rad)) +
    # scale(Day)+
    scale(Temp) +
    scale(StringencyIndex) +
    (1 | genus) + (1 | Species) + (1 | sp_day_year) + (1 | IDLocality) + (1 | sp_loc),
# (1 | Year) + (1 | weekday) + (1 | genus) + (1 | Species) + (1 | sp_day_year) + (1 | IDLocality),
data = s[Country == "Australia"], , REML = FALSE,
        control = lmerControl( 
            optimizer ='optimx', optCtrl=list(method='nlminb')) 
)
est_asx <- est_out(asx, "Australia: (1|genus)+(1|Species)+(1|sp_day_year)+(1|IDLocality)+(1|sp_loc)")
est_asx[, control_for_starting_distance := "no"]

# PL 
pss <- lmer(scale(log(FID)) ~
    scale(Year) +
    scale(log(SD)) +
    scale(log(FlockSize)) +
    scale(log(BodyMass)) +
    scale(sin(rad)) + scale(cos(rad)) +
    # scale(Day)+
    scale(Temp) +
    scale(StringencyIndex) +
    (1 | genus) + (1 | Species) + (1 | sp_day_year),
# (1 | Year) + (1 | weekday) + (1|genus) + (1 | Species) + (1 | sp_day_year) + (1 | IDLocality),
data = s[Country == "Poland"], REML = FALSE
)
est_pss <- est_out(pss, "Poland: (1|genus)+(1|Species)+(1|sp_day_year)")
est_pss[, control_for_starting_distance := "yes"]

psx <- lmer(scale(log(FID)) ~
    scale(Year) +
    # scale(log(SD)) +
    scale(log(FlockSize)) +
    scale(log(BodyMass)) +
    scale(sin(rad)) + scale(cos(rad)) +
    # scale(Day)+
    scale(Temp) +
    scale(StringencyIndex) +
    (1 | genus) + (1 | Species) + (1 | sp_day_year),
# (1 | Year) + (1 | weekday) + (1 | genus) + (1 | Species) + (1 | sp_day_year) + (1 | IDLocality),
data = s[Country == "Poland"], REML = FALSE
)
est_psx <- est_out(psx, "Poland: (1|genus)+(1|Species)+(1|sp_day_year)")
est_psx[, control_for_starting_distance := "no"]

  # combine
    est_mss[, Country := 'All\n(mixed model)']
    est_msx[, Country := "All\n(mixed model)"]
    est_ass[, Country := "Australia"]
    est_asx[, Country := "Australia"]
    est_css[, Country := "Czechia"]
    est_csx[, Country := "Czechia"]
    est_hss[, Country := "Hungary"]
    est_hsx[, Country := "Hungary"]
    est_pss[, Country := "Poland"]
    est_psx[, Country := "Poland"]
    est_fss[, Country := "Finland"]
    est_fsx[, Country := "Finland"]

    os = rbind(est_mss, est_msx, 
            est_ass, est_asx, 
            est_css, est_csx, 
            est_hss, est_hsx,
            est_pss, est_psx, 
            est_fss, est_fsx)
    save(os, file = here::here('Data/dat_est_Stringency_rev.Rdata'))

```
```{r est_str, fig.width=3, fig.height = 2.5}
load(here::here("Data/dat_est_Stringency_rev.Rdata"))
os[predictor %in% c("scale(StringencyIndex)"), predictor := "Stringency Index"]
oso <- os[predictor %in% c("Stringency Index")]
oso[, N:=as.numeric(sub('.*N = ', '', model))]
# add meta-analytical mean
  oso_s = oso[control_for_starting_distance == 'yes']
  met = summary(meta.summaries(d = oso_s$estimate, se = oso_s$sd, method = "fixed", weights = oso_s$N))$summci
  oso_met = data.table(predictor = "Period", estimate = met[2], lwr = met[1], upr = met[3], sd = NA, model = NA, control_for_starting_distance = "yes", Country = "Combined\n(metanalytical)", N = NA)

  oso_sx = oso[control_for_starting_distance == "no"]
  metx = summary(meta.summaries(d = oso_sx$estimate, se = oso_sx$sd, method = "fixed", weights = oso_sx$N))$summci
  oso_metx = data.table(predictor = "Period", estimate = metx[2], lwr = metx[1], upr = metx[3], sd = NA, model = NA, control_for_starting_distance = "no", Country = "Combined\n(metanalytical)", N = NA)
  
  oso = rbind(oso, oso_met, oso_metx)
    
oso[, Country := factor(Country, levels = rev(c("Finland", "Poland", "Czechia", "Hungary", "Australia", "Combined\n(metanalytical)", "All\n(mixed model)")))]

width_ <- .5 # spacing between error bars

#col_ <- c(brewer.pal(n = 12, name = "Paired"), "grey30", "grey80")
#Tol_bright <- c("#EE6677", "#228833", "#4477AA", "#CCBB44", "#66CCEE", "#AA3377", "#BBBBBB")
#Tol_muted <- c("#88CCEE", "#44AA99", "#117733", "#332288", "#DDCC77", "#999933", "#CC6677", "#882255", "#AA4499", "#DDDDDD")
#Tol_light <- c("#BBCC33", "#AAAA00", "#77AADD", "#EE8866", "#EEDD88", "#FFAABB", "#99DDFF", "#44BB99", "#DDDDDD")

# From Color Universal Design (CUD): https://jfly.uni-koeln.de/color/
#Okabe_Ito <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#000000")
#col_ = Okabe_Ito[7:1]
# JAMA and LocusZoom modified order
#col_ =  c("#374E55FF", "#374E55FF", "#DF8F44FF", "#79AF97FF", "#00A1D5FF", "#B24745FF",  "#80796BFF") #"#6A6599FF",
#col_ <- c("#357EBDFF", "#9632B8FF", "#46B8DAFF", "#5CB85CFF", "#EEA236FF", "#D43F3AFF", "#D43F3AFF")[7:1] # "#D43F3AFF", "#B8B8B8FF"
col_ = c("#357EBDFF", "#D43F3AFF", "#46B8DAFF", "#5CB85CFF", "#EEA236FF", "#9632B8FF", "#9632B8FF")[7:1] # "#D43F3AFF", "#B8B8B8FF"
#show_col(col_)

#p = 
ggplot(oso, aes(x = estimate, y = Country, col = Country, shape = control_for_starting_distance)) +
    geom_vline(xintercept = 0, color = "grey", linetype = "dotted") +
    geom_errorbarh(aes(xmin = lwr, xmax = upr), height = 0, position = ggstance::position_dodgev(width_)) +
    # geom_point(position = ggstance::position_dodgev(.6)) +
    geom_point(position = position_dodge(width = width_), bg = "white", size = 1.1) +
    # scale_color_viridis(discrete=TRUE, begin=0, end = 0.5)  +
    # scale_fill_viridis(discrete=TRUE, begin=0, end = 0.5) +
    # geom_text( aes(x = n_pos,label = N), vjust = 0, size = 1.75, position = ggstance::position_dodgev(width_))+ # 3 positions for 3 bars
    # annotate("text", x=log10(3), y=85, label= "Used", col = "grey30", size = 2.5)+

    scale_shape_manual(name = "Controlled for\nstarting distance", guide = guide_legend(reverse = TRUE), values = c(21, 19)) +
    #scale_color_jama(guide = "none")+ #, palette = 'light'
    scale_color_manual(guide = "none", values = col_) + #guide_legend(reverse = TRUE)
    scale_x_continuous(breaks = round(seq(-0.3, 0.4, by = 0.1), 1)) +
    ylab("") +
    xlab("Standardized effect size of Stringency Index\n[on flight initiation distance]") +
    # coord_cartesian(xlim = c(-.15, .15)) +
    # scale_x_continuous(breaks = round(seq(-.15, .15, by = 0.05),2)) +
    theme_bw() +
    theme(
        legend.position = "right",
        legend.title = element_text(size = 7),
        legend.text = element_text(size = 6),
        # legend.spacing.y = unit(0.1, 'cm'),
        legend.key.height = unit(0.5, "line"),
        legend.margin = margin(0, 0, 0, 0),
        # legend.position=c(0.5,1.6),
        plot.title = element_text(color = "grey", size = 7),
        plot.margin = margin(b = 0.5, l = 0.5, t = 0.5, r = 0.5, unit = "pt"),
        panel.grid = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = ax_lines, size = 0.25),
        axis.line.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.ticks.x = element_line(colour = ax_lines, size = 0.25),
        # axis.text.x = element_text()
        axis.ticks.length = unit(1, "pt"),
        axis.text.x = element_text(, size = 6),
        axis.text.y = element_text(colour = "black", size = 7),
        axis.title = element_text(size = 7)
    )

ggsave(here::here("Outputs/Fig_Stringency_rev_width_CustomLocusZoom_v2.png"), width = 8, height = 6, unit = "cm", dpi = 600)
```

**Fig. S_stringency | The effect size with 95%CIs of Stringency Index on flight initiation distance (ln-transformed).**  The general model on all data was controlled for year, starting distance (ln-transformed; filled circles) or not (empty circles), flock size (ln-transformed), temperature (also a proxy for a day within the breeding season: r Pearson = 0.48; Fig. S1) and time of day. To account for circular properties of time, time was transformed into radians (2 × time × π/24) and fitted as sine and cosine of radians (Bulla et al. 2016). All continuous variables were standardised by subtracting the mean and dividing by the standard deviation. The multicollinearity was small as correlations between predictors were  weak (r<XX, Fig. S1). To account for the non-independence of data points (Schielzeth & Forstmeier 2009; Barr et al. 2013), we fitted random intercepts of genus, species, species at a given day and year, site, and species within a site and in the full model Strigency also random slope of Stringency within genus and country. In addition, we attempted to fit Stringency as random slope within genus for country-specific modl, but it explained little variance and estimates remained the same. Note that in the full model, fitting Stringency as random slope at other random intercepts produces similar results (see Fig Sxx). We used this approach with a full dataset with all observations (n = xx), as well as with conservative datasets, one with at least five observations per species  , the other with at least 10 observations per species. Albeit random structure of the full model accounts for the potential country specific effect (country accounted for xx% of variance, Table xx), depicted are also estimates from country specific models, with same predictors and random structure as the full model (excluding the random slopes, and random intercept of country and in case of Poland also site, as specific sites were not noted in Poland). Using meta.summaries function from rmeta R-package we used the country estimates, their standard deviation, and sample size per country to estimate the meta-analytical mean, which reflects the results based on the full mixed effect model. Importantly, the models test for within year changes in esccape distancces due to human changes presence, i.e.  week to week changes in Stringency Index. In other words, the models investigate whether birds flexibly adjust their fear response, a test different to the one intended in this study. Albeit the response of humans to governmental restrictions were likely country specific and presence of humans in parks might have increased in some countries, decreased in others or did not change, the escape distances do not strongly reflect such changes. In other words, the birds are either inflexible or the change in human behavior due to shutdowns was not strong enought, which might have been the case - see Fig. ZZ - decide what to show in this figure (I think 2020-2022 changes in Google Mobility for each country including the fits and refering to supplementary figure). Note however that human presence consistently declined with increased restrictions across countries, albeit the relationships are weak (Fig. 2). 

## predictions for FID ~ Google Mobiliity

```{r }
# predictions
 # full
 mgs <- lmer(scale(log(FID)) ~
     scale(Year) +
     scale(log(SD)) +
     scale(log(FlockSize)) +
     scale(log(BodyMass)) +
     scale(sin(rad)) + scale(cos(rad)) +
     # scale(Day)+
     scale(Temp) +
     scale(parks_percent_change_from_baseline) +
     (scale(parks_percent_change_from_baseline)| genus) + (1 | Species) + (1 | sp_day_year) + 
     (scale(parks_percent_change_from_baseline)| Country) + (1 | IDLocality) + (1 | sp_loc),
 data = ss, REML = FALSE,
 control = lmerControl(
     optimizer = "optimx", optCtrl = list(method = "nlminb")
 )
 )
 est_mgs <- est_out(mgs, "ALL: (scale(Google)|genus) + (1|Species)  + (1|sp_day_year) + (scale(Google)|Country) + (1|IDLocality) +(1|sp_loc)")
 est_mgs[, control_for_starting_distance := "yes"]

 mgx <- lmer(scale(log(FID)) ~
    scale(Year) +
    #scale(log(SD)) +
    scale(log(FlockSize)) +
    scale(log(BodyMass)) +
    scale(sin(rad)) + scale(cos(rad)) +
    # scale(Day)+
    scale(Temp) +
    scale(StringencyIndex) +
    scale(parks_percent_change_from_baseline) +
     (scale(parks_percent_change_from_baseline)| genus) + (1 | Species) + (1 | sp_day_year) + 
     (scale(parks_percent_change_from_baseline) | Country) + (1 | IDLocality) + (1 | sp_loc),
 data = ss, REML = FALSE,
 control = lmerControl(
    optimizer = "optimx", optCtrl = list(method = "nlminb")
)
)
est_mgx <- est_out(mgx, "ALL: (scale(Google)|genus) + (1|Species) + (1|sp_day_year) + (scale(Google)|Country) + (1|IDLocality) +(1|sp_loc)")
est_mgx[, control_for_starting_distance := "no"]


# CZ - singular fits only due to genera estimated as zero (removing it changes no results)
  cgs <- lmer(scale(log(FID)) ~
    #scale(Year) +
    scale(log(SD)) +
    scale(log(FlockSize)) +
    scale(log(BodyMass)) +
    scale(sin(rad)) + scale(cos(rad)) +
    # scale(Day)+
    scale(Temp) +
    scale(parks_percent_change_from_baseline) +
    (1 | genus) + (1| Species) + (1 | sp_day_year) + (1| IDLocality) + (1|sp_loc),
    #(1 | Year) + (1 | weekday) + (1|genus) + (1 | Species) + (1 | sp_day_year) + (1 | IDLocality),
    data = ss[Country == "Czechia"], REML = FALSE
    )
  est_cgs <- est_out(cgs, "Czechia: (1|genus)+(1|Species)+(1|sp_day_year)+(1|IDLocality)+(1|sp_loc)")
  est_cgs[, control_for_starting_distance := "yes"]
  
  cgx <- lmer(scale(log(FID)) ~
    #scale(Year) +
    scale(log(SD)) +
    scale(log(FlockSize)) +
    scale(log(BodyMass)) +
    scale(sin(rad)) + scale(cos(rad)) +
    # scale(Day)+
    scale(Temp) +
    scale(parks_percent_change_from_baseline) +
    (1 | genus) + (1| Species) + (1 | sp_day_year) + (1| IDLocality) + (1|sp_loc),
    #(1 | Year) + (1 | weekday) + (1|genus) + (1 | Species) + (1 | sp_day_year) + (1 | IDLocality),
    data = ss[Country == "Czechia"], REML = FALSE
    )
  est_cgx <- est_out(cgx, "Czechia: (1|genus)+(1|Species)+(1|sp_day_year)+(1|IDLocality)+(1|sp_loc)")
  est_cgx[, control_for_starting_distance := "no"]

# FI
fgs <- lmer(scale(log(FID)) ~
    scale(Year) +
    scale(log(SD)) +
    scale(log(FlockSize)) +
    scale(log(BodyMass)) +
    scale(sin(rad)) + scale(cos(rad)) +
    # scale(Day)+
    scale(Temp) +
    scale(parks_percent_change_from_baseline) +
    (1 | genus) + (1| Species) + (1 | sp_day_year) + (1| IDLocality) + (1|sp_loc),
data = ss[Country == "Finland"], REML = FALSE
)
est_fgs <- est_out(fgs, "Finland: (1|genus)+(1|Species)+(1|sp_day_year)+(scale(StringencyIndex)|IDLocality)+(1|sp_loc)")
est_fgs[, control_for_starting_distance := "yes"]

fgx <- lmer(scale(log(FID)) ~
    scale(Year) +
    # scale(log(SD)) +
    scale(log(FlockSize)) +
    scale(log(BodyMass)) +
    scale(sin(rad)) + scale(cos(rad)) +
    # scale(Day)+
    scale(Temp) +
    scale(parks_percent_change_from_baseline) +
    (1 | genus) + (1| Species) + (1 | sp_day_year) + (1| IDLocality) + (1|sp_loc),
data = ss[Country == "Finland"], REML = FALSE
)
est_fgx <- est_out(fgx, "Finland: (1|genus)+(1|Species)+(1|sp_day_year)+(scale(StringencyIndex)|IDLocality)+(1|sp_loc)")
est_fgx[, control_for_starting_distance := "no"]

# HU - singular fits only due to sp_loc estimated as zero (removing it changes no results)
hgs <- lmer(scale(log(FID)) ~
    scale(Year) +
    scale(log(SD)) +
    scale(log(FlockSize)) +
    scale(log(BodyMass)) +
    scale(sin(rad)) + scale(cos(rad)) +
    # scale(Day)+
    scale(Temp) +
    scale(parks_percent_change_from_baseline) +
     (1 | genus) + (1 | Species) + (1 | sp_day_year) + (1 | IDLocality) + (1 | sp_loc),
# (1 | Year) + (1 | weekday) + (1|genus) + (1 | Species) + (1 | sp_day_year) + (1 | IDLocality),
data = ss[Country == "Hungary"], REML = FALSE
)
est_hgs <- est_out(hgs, "Hungary: (1|genus)+(1|Species)+(1|sp_day_year)+(1|IDLocality)+(1|sp_loc)")
est_hgs[, control_for_starting_distance := "yes"]

hgx <- lmer(scale(log(FID)) ~
    scale(Year) +
    # scale(log(SD)) +
    scale(log(FlockSize)) +
    scale(log(BodyMass)) +
    scale(sin(rad)) + scale(cos(rad)) +
    # scale(Day)+
    scale(Temp) +
    scale(parks_percent_change_from_baseline) +
    (1 | genus) + (1 | Species) + (1 | sp_day_year) + (1 | IDLocality) + (1 | sp_loc),
# (1 | Year) + (1 | weekday) + (1 | genus) + (1 | Species) + (1 | sp_day_year) + (1 | IDLocality),
data = ss[Country == "Hungary"], REML = FALSE
)
est_hgx <- est_out(hgx, "Hungary: (1|genus)+(1|Species)+(1|sp_day_year)+(1|IDLocality)+(1|sp_loc)")
est_hgx[, control_for_starting_distance := "no"]

# AU - singular fits only due to Year and random slope estimated as zero (removing those changes no results)
ags <- lmer(scale(log(FID)) ~
    scale(Year) +
    scale(log(SD)) +
    scale(log(FlockSize)) +
    scale(log(BodyMass)) +
    scale(sin(rad)) + scale(cos(rad)) +
    # scale(Day)+
    scale(Temp) +
    scale(parks_percent_change_from_baseline) +
      (1 | genus) + (1 | Species) + (1 | sp_day_year) + (1 | IDLocality) + (1 | sp_loc),
# (1 | Year) + (1 | weekday) + (1|genus) + (1 | Species) + (1 | sp_day_year) + (1 | IDLocality),
data = ss[Country == "Australia"], REML = FALSE
)
est_ags <- est_out(ags, "Australia: (1|genus)+(1|Species)+(1|sp_day_year)+(1|IDLocality)+(1|sp_loc)")
est_ags[, control_for_starting_distance := "yes"]

agx <- lmer(scale(log(FID)) ~
    scale(Year) +
    # scale(log(SD)) +
    scale(log(FlockSize)) +
    scale(log(BodyMass)) +
    scale(sin(rad)) + scale(cos(rad)) +
    # scale(Day)+
    scale(Temp) +
    scale(parks_percent_change_from_baseline) +
    (1 | genus) + (1 | Species) + (1 | sp_day_year) + (1 | IDLocality) + (1 | sp_loc),
# (1 | Year) + (1 | weekday) + (1 | genus) + (1 | Species) + (1 | sp_day_year) + (1 | IDLocality),
data = ss[Country == "Australia"], , REML = FALSE,
        control = lmerControl( 
            optimizer ='optimx', optCtrl=list(method='nlminb')) 
)
est_agx <- est_out(agx, "Australia: (1|genus)+(1|Species)+(1|sp_day_year)+(1|IDLocality)+(1|sp_loc)")
est_agx[, control_for_starting_distance := "no"]

# PL 
pgs <- lmer(scale(log(FID)) ~
    scale(Year) +
    scale(log(SD)) +
    scale(log(FlockSize)) +
    scale(log(BodyMass)) +
    scale(sin(rad)) + scale(cos(rad)) +
    # scale(Day)+
    scale(Temp) +
    scale(parks_percent_change_from_baseline) +
    (1 | genus) + (1 | Species) + (1 | sp_day_year),
# (1 | Year) + (1 | weekday) + (1|genus) + (1 | Species) + (1 | sp_day_year) + (1 | IDLocality),
data = ss[Country == "Poland"], REML = FALSE
)
est_pgs <- est_out(pgs, "Poland: (1|genus)+(1|Species)+(1|sp_day_year)")
est_pgs[, control_for_starting_distance := "yes"]

pgx <- lmer(scale(log(FID)) ~
    scale(Year) +
    # scale(log(SD)) +
    scale(log(FlockSize)) +
    scale(log(BodyMass)) +
    scale(sin(rad)) + scale(cos(rad)) +
    # scale(Day)+
    scale(Temp) +
    scale(parks_percent_change_from_baseline) +
    (1 | genus) + (1 | Species)+ (1 | sp_day_year),
# (1 | Year) + (1 | weekday) + (1 | genus) + (1 | Species) + (1 | sp_day_year) + (1 | IDLocality),
data = ss[Country == "Poland"], REML = FALSE
)
est_pgx <- est_out(pgx, "Poland: (1|genus)+(1|Species)+(1|sp_day_year)")
est_pgx[, control_for_starting_distance := "no"]

  # combine
    est_mgs[, Country := 'All\n(mixed model)']
    est_mgx[, Country := "All\n(mixed model)"]
    est_ags[, Country := "Australia"]
    est_agx[, Country := "Australia"]
    est_cgs[, Country := "Czechia"]
    est_cgx[, Country := "Czechia"]
    est_hgs[, Country := "Hungary"]
    est_hgx[, Country := "Hungary"]
    est_pgs[, Country := "Poland"]
    est_pgx[, Country := "Poland"]
    est_fgs[, Country := "Finland"]
    est_fgx[, Country := "Finland"]

    og = rbind(est_mgs, est_mgx, 
            est_ags, est_agx, 
            est_cgs, est_cgx, 
            est_hgs, est_hgx,
            est_pgs, est_pgx, 
            est_fgs, est_fgx)
    save(og, file = here::here('Data/dat_est_Google_rev.Rdata'))

```
```{r est_gm, fig.width=3, fig.height = 2.5}
load(here::here("Data/dat_est_Google_rev.Rdata"))
og[predictor %in% c("scale(parks_percent_change_from_baseline)"), predictor := "Google Mobility"]
ogo <- og[predictor %in% c("Google Mobility")]
ogo[, N:=as.numeric(sub('.*N = ', '', model))]
# add meta-analytical mean
  ogo_s = ogo[control_for_starting_distance == 'yes']
  met = summary(meta.summaries(d = ogo_s$estimate, se = ogo_s$sd, method = "fixed", weights = ogo_s$N))$summci
  ogo_met = data.table(predictor = "Period", estimate = met[2], lwr = met[1], upr = met[3], sd = NA, model = NA, control_for_starting_distance = "yes", Country = "Combined\n(metanalytical)", N = NA)

  ogo_sx = ogo[control_for_starting_distance == "no"]
  metx = summary(meta.summaries(d = ogo_sx$estimate, se = ogo_sx$sd, method = "fixed", weights = ogo_sx$N))$summci
  ogo_metx = data.table(predictor = "Period", estimate = metx[2], lwr = metx[1], upr = metx[3], sd = NA, model = NA, control_for_starting_distance = "no", Country = "Combined\n(metanalytical)", N = NA)
  
  ogo = rbind(ogo, ogo_met, ogo_metx)
    
ogo[, Country := factor(Country, levels = rev(c("Finland", "Poland", "Czechia", "Hungary", "Australia", "Combined\n(metanalytical)", "All\n(mixed model)")))]

width_ <- .5 # spacing between error bars

#col_ <- c(brewer.pal(n = 12, name = "Paired"), "grey30", "grey80")
#Tol_bright <- c("#EE6677", "#228833", "#4477AA", "#CCBB44", "#66CCEE", "#AA3377", "#BBBBBB")
#Tol_muted <- c("#88CCEE", "#44AA99", "#117733", "#332288", "#DDCC77", "#999933", "#CC6677", "#882255", "#AA4499", "#DDDDDD")
#Tol_light <- c("#BBCC33", "#AAAA00", "#77AADD", "#EE8866", "#EEDD88", "#FFAABB", "#99DDFF", "#44BB99", "#DDDDDD")

# From Color Universal Design (CUD): https://jfly.uni-koeln.de/color/
#Okabe_Ito <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#000000")
#col_ = Okabe_Ito[7:1]
# JAMA and LocusZoom modified order
#col_ =  c("#374E55FF", "#374E55FF", "#DF8F44FF", "#79AF97FF", "#00A1D5FF", "#B24745FF",  "#80796BFF") #"#6A6599FF",
#col_ <- c("#357EBDFF", "#9632B8FF", "#46B8DAFF", "#5CB85CFF", "#EEA236FF", "#D43F3AFF", "#D43F3AFF")[7:1] # "#D43F3AFF", "#B8B8B8FF"
col_ = c("#357EBDFF", "#D43F3AFF", "#46B8DAFF", "#5CB85CFF", "#EEA236FF", "#9632B8FF", "#9632B8FF")[7:1] # "#D43F3AFF", "#B8B8B8FF"
#show_col(col_)

#p = 
ggplot(ogo, aes(x = estimate, y = Country, col = Country, shape = control_for_starting_distance)) +
    geom_vline(xintercept = 0, color = "grey", linetype = "dotted") +
    geom_errorbarh(aes(xmin = lwr, xmax = upr), height = 0, position = ggstance::position_dodgev(width_)) +
    # geom_point(position = ggstance::position_dodgev(.6)) +
    geom_point(position = position_dodge(width = width_), bg = "white", size = 1.1) +
    # scale_color_viridis(discrete=TRUE, begin=0, end = 0.5)  +
    # scale_fill_viridis(discrete=TRUE, begin=0, end = 0.5) +
    # geom_text( aes(x = n_pos,label = N), vjust = 0, size = 1.75, position = ggstance::position_dodgev(width_))+ # 3 positions for 3 bars
    # annotate("text", x=log10(3), y=85, label= "Used", col = "grey30", size = 2.5)+

    scale_shape_manual(name = "Controlled for\nstarting distance", guide = guide_legend(reverse = TRUE), values = c(21, 19)) +
    #scale_color_jama(guide = "none")+ #, palette = 'light'
    scale_color_manual(guide = "none", values = col_) + #guide_legend(reverse = TRUE)
    scale_x_continuous(breaks = round(seq(-0.3, 0.2, by = 0.1), 1)) +
    ylab("") +
    xlab("Standardized effect size of Google Mobility (human presence)\n[on flight initiation distance]") +
    # coord_cartesian(xlim = c(-.15, .15)) +
    # scale_x_continuous(breaks = round(seq(-.15, .15, by = 0.05),2)) +
    theme_bw() +
    theme(
        legend.position = "right",
        legend.title = element_text(size = 7),
        legend.text = element_text(size = 6),
        # legend.spacing.y = unit(0.1, 'cm'),
        legend.key.height = unit(0.5, "line"),
        legend.margin = margin(0, 0, 0, 0),
        # legend.position=c(0.5,1.6),
        plot.title = element_text(color = "grey", size = 7),
        plot.margin = margin(b = 0.5, l = 0.5, t = 0.5, r = 0.5, unit = "pt"),
        panel.grid = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = ax_lines, size = 0.25),
        axis.line.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.ticks.x = element_line(colour = ax_lines, size = 0.25),
        # axis.text.x = element_text()
        axis.ticks.length = unit(1, "pt"),
        axis.text.x = element_text(, size = 6),
        axis.text.y = element_text(colour = "black", size = 7),
        axis.title = element_text(size = 7)
    )

ggsave(here::here("Outputs/Fig_Google_rev_width_CustomLocusZoom_v2.png"), width = 8, height = 6, unit = "cm", dpi = 600)
```

**Fig. S_Google | The effect size with 95%CIs of Google Mobility on flight initiation distance (ln-transformed).**  The general model on all data was controlled for year, starting distance (ln-transformed; filled circles) or not (empty circles), flock size (ln-transformed), temperature (also a proxy for a day within the breeding season: r Pearson = 0.48; Fig. S1) and time of day. To account for circular properties of time, time was transformed into radians (2 × time × π/24) and fitted as sine and cosine of radians (Bulla et al. 2016). All continuous variables were standardised by subtracting the mean and dividing by the standard deviation. The multicollinearity was small as correlations between predictors were  weak (r<XX, Fig. S1). To account for the non-independence of data points (Schielzeth & Forstmeier 2009; Barr et al. 2013), we fitted random intercepts of genus, species, species at a given day and year, site, and species within a site and in the full model also random slope of Google Mobility within genus and country.  Note that in the full model, fitting Google Mobility as random slope at other random intercepts produces similar results (see Fig Sxx). We used this approach with a full dataset with all observations (n = xx), as well as with conservative datasets, one with at least five observations per species  , the other with at least 10 observations per species. Albeit random structure of the full model accounts for the potential country specific effect (country accounted for xx% of variance, Table xx), depicted are also estimates from country specific models, with same predictors and random structure as the full model (excluding the random slopes and also random intercepts of country and in case of Poland  site, as specific sites were not noted in Poland). Using meta.summaries function from rmeta R-package we used the country estimates, their standard deviation, and sample size per country to estimate the meta-analytical mean, which reflects the results based on the full mixed effect model. Importantly, the models test for within year changes in esccape distancces due to changes in human presence, i.e.  day to day changes in Google Mobility.  In other words, the models investigate whether birds flexibly adjust their fear response on a day to day basis (regardless of COVID-19 restrictions), a test different to the one intended in this study. The effect sizes are small, mostly close to zero, and hihgly uncertain. 


## Genus and spcies 'raw' data visualisations

```{r }
# prepare data for Figure 2, S3, S5
dxx <- d[paste(IDLocality, Species) %in% paste(pp$IDLocality, pp$Species)]
# length(dxx[, unique(paste(sp_loc, Covid))])
m <- lm(log(FID) ~ log(SD), dxx)
dxx[, resid_FID := resid(m)]
a <- dxx[, .(mean(resid_FID), sd(resid_FID), mean(FID), .N), by = .(Country, IDLocality, genus, Species, sp_loc, Covid)]
setnames(a, old = c("V1", "V2", "V3"), new = c("resid_FID_avg", "SD", "FID_avg"))
a[is.na(SD), SD := 0]

aw <- reshape(a, idvar = c("Country", "IDLocality", "genus", "Species", "sp_loc"), timevar = "Covid", direction = "wide")
aw[, Species := gsub("[_]", " ", Species)]
aw <- merge(aw, t, all.x = TRUE)
table(aw$Family)

x <- aw[, .N, by = Species]
x[order(Species)]
aw[, genus2 := genus]
aw[Species %in% x[N %in% c(1, 2), Species], genus2 := "other"]

aw[genus2 == "Phoenicurus", unique(Species)]

ph[genus2 == "Motacilla" | uid %in% c("67a9ecfd-58ba-44a4-9986-243b6e610419"), uid := "cf522e02-35cc-44f5-841c-0e642987c2e4"]
ph[genus2 == "Sylvia", uid := "67a9ecfd-58ba-44a4-9986-243b6e610419"]

ph[, size := 0.2]
ph[genus2 %in% c("Anas", "Columba", "Dendrocopos", "Sturnus"), size := c(0.25, 0.25, 0.15, 0.1)]
ph[, FID_avg.0 := 1.5]
ph[, FID_avg.1 := 20]
ph[genus2 %in% c("Anas", "Columba"), FID_avg.0 := c(1.7, 1.7)]

ph[, resid_FID_avg.0 := -1.7]
ph[, resid_FID_avg.1 := 0.7]
ph[genus2 %in% c("Anas", "Columba"), resid_FID_avg.0 := c(-1.6, -1.6)]

ph[, genus2 := factor(genus2, levels = c("Anas", "Larus", "Columba", "Dendrocopos", "Picus", "Motacilla", "Erithacus", "Phoenicurus", "Turdus", "Sylvia", "Parus", "Sitta", "Pica", "Garrulus", "Corvus", "Sturnus", "Passer", "Fringilla", "other"))]

aw[, genus2 := factor(genus2, levels = c("Anas", "Larus", "Columba", "Dendrocopos", "Picus", "Motacilla", "Erithacus", "Phoenicurus", "Turdus", "Sylvia", "Parus", "Sitta", "Pica", "Garrulus", "Corvus", "Sturnus", "Passer", "Fringilla", "other"))]

aw[, genus2 := factor(genus2, levels = c("Anas", "Larus", "Columba", "Dendrocopos", "Picus", "Motacilla", "Erithacus", "Phoenicurus", "Turdus", "Sylvia", "Parus", "Sitta", "Pica", "Garrulus", "Corvus", "Sturnus", "Passer", "Fringilla", "other"))]

```
```{r genus, fig.width=8, fig.height = 8}
# Fig 2 & left panel of S5- plot from files
anas <- rasterGrob(change_col("#CCCCCC", readPNG("Data/Pics/Anas.png")))
columba <- rasterGrob(change_col("#CCCCCC", readPNG("Data/Pics/Columba.png")))
Dendrocopos <- rasterGrob(change_col("#CCCCCC", readPNG("Data/Pics/Dendrocopos.png")))
Larus <- rasterGrob(change_col("#CCCCCC", readPNG("Data/Pics/Larus_flip.png")))
Picus <- rasterGrob(change_col("#CCCCCC", readPNG("Data/Pics/Picus.png")))
Motacilla <- rasterGrob(change_col("#CCCCCC", readPNG("Data/Pics/Motacilla.png")))
Erithacus <- rasterGrob(change_col("#CCCCCC", readPNG("Data/Pics/Erithacus.png")))
Phoenicurus <- rasterGrob(change_col("#CCCCCC", readPNG("Data/Pics/Phoenicurus.png")))
Turdus <- rasterGrob(change_col("#CCCCCC", readPNG("Data/Pics/Turdus.png")))
Sylvia <- rasterGrob(change_col("#CCCCCC", readPNG("Data/Pics/Sylvia.png")))
Parus <- rasterGrob(change_col("#CCCCCC", readPNG("Data/Pics/Parus_flip.png")))
Sitta <- rasterGrob(change_col("#CCCCCC", readPNG("Data/Pics/Sitta.png")))
Pica <- rasterGrob(change_col("#CCCCCC", readPNG("Data/Pics/Pica.png")))
Garrulus <- rasterGrob(change_col("#CCCCCC", readPNG("Data/Pics/Garrulus.png")))
Corvus <- rasterGrob(change_col("#CCCCCC", readPNG("Data/Pics/Corvus.png")))
Sturnus <- rasterGrob(change_col("#CCCCCC", readPNG("Data/Pics/Sturnus.png")))
Passer <- rasterGrob(change_col("#CCCCCC", readPNG("Data/Pics/Passer_flip.png")))
Fringilla <- rasterGrob(change_col("#CCCCCC", readPNG("Data/Pics/Fringilla.png")))
other <- rasterGrob(change_col("#CCCCCC", readPNG("Data/Pics/other_flip.png")))

ann_text <- data.frame(
  FID_avg.0 = 8, FID_avg.1 = 10, lab = "Text",
  genus2 = factor("Anas", levels = c("Anas", "Larus", "Columba", "Dendrocopos", "Picus", "Motacilla", "Erithacus", "Phoenicurus", "Turdus", "Sylvia", "Parus", "Sitta", "Pica", "Garrulus", "Corvus", "Sturnus", "Passer", "Fringilla", "other"))
)
ann_text2 <- data.frame(
  FID_avg.0 = 6, FID_avg.1 = 3, lab = "Text",
  genus2 = factor("Larus", levels = c("Anas", "Larus", "Columba", "Dendrocopos", "Picus", "Motacilla", "Erithacus", "Phoenicurus", "Turdus", "Sylvia", "Parus", "Sitta", "Pica", "Garrulus", "Corvus", "Sturnus", "Passer", "Fringilla", "other"))
)

aw2 <- data.frame(FID_avg.0 = c(11.25, 11.25), FID_avg.1 = c(3.5, 5.8), genus2 = factor("Larus", levels = c("Anas", "Larus", "Columba", "Dendrocopos", "Picus", "Motacilla", "Erithacus", "Phoenicurus", "Turdus", "Sylvia", "Parus", "Sitta", "Pica", "Garrulus", "Corvus", "Sturnus", "Passer", "Fringilla", "other")))

col3_ <- c("#357EBDFF", "#D43F3AFF", "#46B8DAFF", "#5CB85CFF", "#EEA236FF", "#9632B8FF", "#9632B8FF")[7:1]
col3__ <- col3_[3:7]

aw[, Country := factor(Country, levels = rev(c("Finland", "Poland", "Czechia", "Hungary", "Australia")))]

g <-
  ggplot(aw, aes(x = FID_avg.0, y = FID_avg.1)) +
  # geom_errorbar(aes(ymin = FID_avg.1-SD.1, ymax = FID_avg.1+SD.1, col = Country), width = 0) +
  # geom_errorbar(aes(xmin = FID_avg.0-SD.0, xmax = FID_avg.0+SD.0, col = Country), width = 0) +
  # geom_point(pch = 21, alpha = 0.7, aes(col = Country)) +
  annotation_custom2(anas, data = ph[genus2 == "Anas"], xmin = 0.05, xmax = 0.5, ymax = 2.6) +
  annotation_custom2(Larus, data = ph[genus2 == "Larus"], xmin = 0.05, xmax = 0.5, ymax = 2.6) +
  annotation_custom2(columba, data = ph[genus2 == "Columba"], xmin = 0.05, xmax = 0.4, ymax = 2.7) +
  annotation_custom2(Dendrocopos, data = ph[genus2 == "Dendrocopos"], xmin = 0.05, xmax = 0.25, ymax = 2.6) +
  annotation_custom2(Picus, data = ph[genus2 == "Picus"], xmin = 0.05, xmax = 0.4, ymax = 2.7) +
  annotation_custom2(Motacilla, data = ph[genus2 == "Motacilla"], xmin = 0.05, xmax = 0.5, ymax = 2.7) +
  annotation_custom2(Erithacus, data = ph[genus2 == "Erithacus"], xmin = 0.05, xmax = 0.35, ymax = 2.7) +
  annotation_custom2(Phoenicurus, data = ph[genus2 == "Phoenicurus"], xmin = 0.05, xmax = 0.35, ymax = 2.7) +
  annotation_custom2(Turdus, data = ph[genus2 == "Turdus"], xmin = 0.05, xmax = 0.5, ymax = 2.7) +
  annotation_custom2(Sylvia, data = ph[genus2 == "Sylvia"], xmin = 0.05, xmax = 0.5, ymax = 2.7) +
  annotation_custom2(Parus, data = ph[genus2 == "Parus"], xmin = 0.05, xmax = 0.42, ymax = 2.7) +
  annotation_custom2(Sitta, data = ph[genus2 == "Sitta"], xmin = 0.05, xmax = 0.5, ymax = 2.7) +
  annotation_custom2(Pica, data = ph[genus2 == "Pica"], xmin = 0.05, xmax = 0.5, ymax = 2.5) +
  annotation_custom2(Garrulus, data = ph[genus2 == "Garrulus"], xmin = 0.05, xmax = 0.6, ymax = 2.7) +
  annotation_custom2(Corvus, data = ph[genus2 == "Corvus"], xmin = 0.05, xmax = 0.4, ymax = 2.55) +
  annotation_custom2(Sturnus, data = ph[genus2 == "Sturnus"], xmin = 0.05, xmax = 0.24, ymax = 2.65) +
  annotation_custom2(Passer, data = ph[genus2 == "Passer"], xmin = 0.05, xmax = 0.36, ymax = 2.65) +
  annotation_custom2(Fringilla, data = ph[genus2 == "Fringilla"], xmin = 0.05, xmax = 0.5, ymax = 2.7) +
  annotation_custom2(other, data = ph[genus2 == "other"], xmin = 0.05, xmax = 0.45, ymax = 2.4) +
  geom_point(pch = 21, alpha = 0.7, aes(fill = Country), col = 'grey20', alpha =0.8)+#col = "white") +
  # ggtitle ("Sim based")+
  geom_abline(intercept = 0, slope = 1, lty = 3, col = "grey80") +
  geom_line(data = aw2, col = "grey80", lwd = 0.25) +
  geom_text(data = ann_text, label = "No difference", col = "grey80", angle = 45, size = 2) +
  geom_text(data = ann_text2, label = "Species mean / site", col = "grey60", size = 2, ) +
  facet_wrap(~genus2) +
  # geom_phylopic(data = o, aes(image = uid),  color = "grey80", size = o$size) + # ,
  #scale_fill_viridis(discrete = TRUE, guide = guide_legend(reverse = FALSE)) +
  scale_fill_manual(values = col3__,guide = guide_legend(reverse = TRUE)) +
  scale_x_continuous("Before COVID-19 shutdown - flight initiation distance [m]", expand = c(0, 0), trans = "log10") +
  scale_y_continuous("During COVID-19 shutdown - flight initiation distance [m]", expand = c(0, 0), trans = "log10") +
  # labs(title = "Species means per sampling location")+
  theme_MB +
  theme(
    plot.title = element_text(size = 7),
    strip.background = element_blank(),
    # panel.spacing = unit(1, "mm"),
    legend.position = c(1, 0.025),
    legend.justification = c(1, -0.05)
  )
gg <- ggplotGrob(g) # gg$layout$name
ggx <- gtable_filter_remove(gg,
  name = paste0("axis-b-", c(2, 4), "-4"),
  trim = FALSE
)
# grid.draw(ggx)
# ggsave('Outputs/Fig_2_width-114mm.png',ggx, width=4.5,height=4.5,dpi=600) # 11.43cm # with label on top
ggsave(here::here("Outputs/Fig_2_width-122mm_col_grey_rev.png"), ggx, width = 4.8, height = 4.5, dpi = 600) # 12.2cm # with label inside

```
```{r genus_compare, fig.width=8*2, fig.height = 8}
   # Fig S5 right panel
   g2 <-
     ggplot(aw, aes(x = resid_FID_avg.0, y = resid_FID_avg.1)) +
     geom_point(pch = 21, alpha = 0.7, aes(fill = Country), col = 'grey20', alpha =0.8)+#col = "white") +
     geom_abline(intercept = 0, slope = 1, lty = 3, col = "grey80") +
     facet_wrap(~genus2) +
     # geom_phylopic(data = o, aes(image = uid),  color = "grey80", size = o$size) +
     #scale_fill_viridis(discrete = TRUE, guide = guide_legend(reverse = TRUE)) +
     scale_fill_manual(values = col3__,guide = guide_legend(reverse = TRUE)) +
     scale_x_continuous("Before COVID-19 shutdown - residual escape distance", expand = c(0, 0)) +
     scale_y_continuous("During COVID-19 shutdown - residual escape distance", expand = c(0, 0)) +
     # labs(title = "Species means per sampling location")+
     theme_MB +
     theme(
       plot.title = element_text(size = 7),
       strip.background = element_blank(),
       legend.position = "none",
       # legend.position = c(1, 0),
       legend.justification = c(1, 0)
     )

   gg2 <- ggplotGrob(g2) # gg$layout$name
   ggx2 <- gtable_filter_remove(gg2,
     name = paste0("axis-b-", c(2, 4), "-4"),
     trim = FALSE
   )
   # Fig S5 combine
   grid.draw(cbind(cbind(ggx, ggx2, size = "last")))
   ggsave(here::here("Outputs/Fig_S5_rev_v2.png"), cbind(ggx, ggx2, size = "last"), width = 4.8 * 2, height = 4.5, dpi = 600)

   # Fig S5 legend
   # correlation between mean fid and residual fid
   ggplot(a, aes(x = resid_FID_avg)) +
     geom_histogram()
   ggplot(a, aes(x = FID_avg, y = resid_FID_avg)) +
     stat_smooth() +
     geom_point() +
     stat_cor(aes(label = ..r.label..), label.x = 0.3, size = 2) +
     scale_x_continuous(trans = "log")
   ggplot(dxx, aes(x = FID, y = resid_FID)) +
     stat_smooth() +
     geom_point() +
     stat_cor(aes(label = ..r.label..), label.x = 0.3, size = 2) +
     scale_x_continuous(trans = "log")
   cor(log(dxx$FID), dxx$resid_FID)
   cor(log(a$FID_avg), a$resid_FID_avg)
   ggplot(aw, aes(x = N.0 - N.1)) +
     geom_histogram()
   nrow(aw[abs(N.0 - N.1) > 2])
   nrow(aw[!abs(N.0 - N.1) > 2])

# Fig S3 - species means
```
```{r species_compare, fig.width=5.3, fig.height = 7}
      aw[, sp2 := gsub(" ", "\n", Species)]
      ann_text <- data.frame(FID_avg.0 = 8, FID_avg.1 = 10,lab = "Text",
                       Species = factor('Aegithalos caudatus',levels = levels(as.factor(aw$Species))))
      ann_text$sp2 = gsub(" ", "\n", ann_text$Species)
      g3 = 
        ggplot(aw, aes(x = FID_avg.0, y = FID_avg.1)) + 
          #geom_errorbar(aes(ymin = FID_avg.1-SD.1, ymax = FID_avg.1+SD.1, col = Country), width = 0) +
          #geom_errorbar(aes(xmin = FID_avg.0-SD.0, xmax = FID_avg.0+SD.0, col = Country), width = 0) +
          #geom_point(pch = 21, alpha = 0.7, aes(col = Country)) + 
          geom_point(pch = 21, alpha = 0.7, aes(fill = Country), col = 'grey20', alpha =0.8)+#col = "white") +
            #ggtitle ("Sim based")+
          geom_abline(intercept = 0, slope = 1, lty =3, col = "grey80")+
          geom_text(data = ann_text,label = "No difference", col = "grey80",angle = 45, size = 2) + 
          facet_wrap(~sp2) +
          #geom_phylopic(data = o, aes(image = uid),  color = "grey80", size = o$size) + # ,
          #scale_fill_viridis(discrete=TRUE,guide = guide_legend(reverse = FALSE))  +
          scale_fill_manual(values = col3__,guide = guide_legend(reverse = TRUE)) +
          scale_x_continuous("Before COVID-19 shutdown - flight initiation distance [m]", expand = c(0, 0), trans = 'log10') +
          scale_y_continuous("During COVID-19 shutdown - flight initiation distance [m]", expand = c(0, 0), trans = 'log10') +
          labs(title = "Species means per sampling location")+
          theme_MB  +
          theme(
                  plot.title = element_text(size=7),
                  strip.background = element_blank(),
                  #panel.spacing = unit(1, "mm"),
                  legend.position = c(0.96, 0.0),
                  legend.justification = c(1, 0)
                  )  
        gg3 <- ggplotGrob(g3) #gg2$layout$name
        ggx3 <- gtable_filter_remove(gg3, name = c(paste0("axis-b-", c(2, 4), "-7"), "axis-b-6-6"),
                                         trim = FALSE)
        #grid.draw(ggx2)
        ggsave(here::here('Outputs/Fig_S3_species_rev_v2.png'),ggx3, width=13.5,height=17.5,unit = 'cm', dpi=600) # 11.43cm
        gg3
# Figure S7 - stringency trends
```
```{r species_string, fig.width=6, fig.height = 7.5}
ss <- s[Nsp > 9]
ss[, sp2 := gsub(" ", "\n", sp)]
ss[, Country := factor(Country, levels = rev(c("Finland", "Poland", "Czechia", "Hungary", "Australia")))]
g2 <-
  ggplot(ss, aes(x = StringencyIndex, y = FID)) +
  # stat_smooth(method = 'rlm', se = FALSE, col = 'black', lwd = 0.5)+
  geom_point(pch = 21, alpha = 0.7, aes(fill = Country), col = 'grey20', alpha =0.8)+
  stat_smooth(se = FALSE, aes(colour = "Locally weighted\nsmoothing"), lwd = 0.5) + # show_guide=TRUE
  facet_wrap(~sp2, ncol = 6) +
  scale_fill_manual(values = col3__,guide = guide_legend(reverse = TRUE)) +
  #scale_fill_viridis(discrete = TRUE, guide = guide_legend(reverse = FALSE)) +
  scale_x_continuous("Stringency index of governmental COVID-19 restrictions", expand = c(0, 0)) +
  scale_y_continuous("Flight initiation distance [m]", expand = c(0, 0), trans = "log10") +
  # annotate("text", x = 1, y = 1, label = c(rep("", 52),"Observation"), hjust = -0.08, size = 1) +
  # labs(title = "Species means per sampling location")+
  scale_colour_manual(values = c("grey60")) +
  # scale_color_manual(name = 'try', values = c('LOESS smoothed = "grey60"'))+
  theme_MB +
  theme(
    plot.title = element_text(size = 7),
    strip.background = element_blank(),
    # strip.text.x = element_text(size = 4.5, color="grey30",  margin=margin(1,1,1,1,"mm")),
    # panel.spacing = unit(1, "mm"),
    legend.position = c(1, 0.01),
    legend.justification = c(1, 0),
    legend.title = element_blank(),
    # legend.spacing.y = unit(-0.78, "cm")
    # legend.spacing.y = unit(0.02, "cm") use if LOESS smooth text as legend
    legend.spacing.y = unit(-0.9, "cm")
  )

gg2 <- ggplotGrob(g2) # gg$layout$name
ggx2 <- gtable_filter_remove(gg2, name = paste0("axis-b-", c(2, 4), "-9"), trim = FALSE)
# grid.draw(ggx)
ggsave(here::here("Outputs/Fig_3_width-152mm_2-row_v2.png"), ggx2, width = 15.24, height = 19, unit = "cm", dpi = 600)

# Fig S8 - google trrends
```
```{r species_google, fig.width=6, fig.height = 7.5}
ss[, NspC := .N, by = "sp_country"]
ssc <- ss[NspC > 9]
ssc[, sp2 := gsub(" ", "\n", sp)]
ssc[, Google_mobility := parks_percent_change_from_baseline]
ssc[, Country := factor(Country, levels = rev(c("Finland", "Poland", "Czechia", "Hungary", "Australia")))]

# two rows labels
g2 <-
  ggplot(ssc, aes(x = Google_mobility, y = FID, )) +
  # stat_smooth(method = 'rlm', se = FALSE, col = 'black', lwd = 0.5)+
  geom_point(pch = 21, alpha = 0.7, aes(fill = Country), col = 'grey20', alpha =0.8)+
  stat_smooth(se = FALSE, aes(colour = "Locally weighted\nsmoothing"), lwd = 0.5) + # show_guide=TRUE
  facet_wrap(~sp2, ncol = 6) +
  scale_fill_manual(values = col3__,guide = guide_legend(reverse = TRUE)) +
   scale_colour_manual(values = c("grey60")) +
  scale_x_continuous("Google Mobility", expand = c(0, 0)) +
  scale_y_continuous("Flight initiation distance [m]", expand = c(0, 0), trans = "log10") +

  # annotate("text", x = 1, y = 1, label = c(rep("", 52),"Observation"), hjust = -0.08, size = 1) +
  # labs(title = "Species means per sampling location")+
  # scale_colour_manual(values=c('grey60'))+
  # scale_color_manual(name = 'try', values = c('LOESS smoothed = "grey60"'))+
  theme_MB +
  theme(
    plot.title = element_text(size = 7),
    strip.background = element_blank(),
    # strip.text.x = element_text(size = 4.5, color="grey30",  margin=margin(1,1,1,1,"mm")),
    # panel.spacing = unit(1, "mm"),
    legend.position = c(1, 0.01),
    legend.justification = c(1, 0),
    legend.title = element_blank(),
    # legend.spacing.y = unit(-0.78, "cm")
    # legend.spacing.y = unit(0.02, "cm") use if LOESS smooth text as legend
    legend.spacing.y = unit(-0.9, "cm")
  )

gg2 <- ggplotGrob(g2) # gg$layout$name
ggx2 <- gtable_filter_remove(gg2, name = c("axis-b-2-9", "axis-b-5-8"), trim = FALSE) # paste0("axis-b-", c(2, 4), "-9")
grid.draw(ggx2)
ggsave(here::here("Outputs/Fig_S8_width-152mm_Google.png"), ggx2, width = 15.24, height = 19, unit = "cm", dpi = 600)
```

## alternative models 

```{r }
  # prepare estimates Period
    # 01a all data, main text
      m1a <- lmer(scale(log(FID)) ~
        scale(log(SD)) +
        scale(log(FlockSize)) +
        scale(log(BodyMass)) +
        scale(sin(rad)) + scale(cos(rad)) +
        # scale(Day)+
        scale(Temp) +
        scale(Covid) +
        (1 | Year) + (1 | weekday) + (1 | genus) + (1 | Species) + (1 | sp_day_year) + (scale(Covid) | Country) + (scale(Covid) | IDLocality) + (1 | sp_loc),
      data = d, REML = FALSE,
       control = lmerControl(
         optimizer = "optimx", optCtrl = list(method = "nlminb")
       )
      )
     est_m1a = est_out(m1a, "01a) (1|Year) + (1|weekday) + (1|genus) + (1|Species) + (1|sp_day_year) + (scale(Covid)|Country) + (scale(Covid)|IDLocality) + (1|sp_loc)")
    # 01b all data, all random slopes - singularity 
      m1b=lmer(scale(log(FID))~
                  scale(log(SD))+
                  scale(log(FlockSize))+
                  scale(log(BodyMass))+
                  scale(sin(rad)) + scale(cos(rad)) + 
                  #scale(Day)+
                  scale(Temp)+
                  scale(Covid)+
                  (1|Year) + (1 | weekday) + (scale(Covid)|genus)+(scale(Covid)|Species)+(1|sp_day_year) + (scale(Covid)|Country) +(scale(Covid)|IDLocality) + (scale(Covid)|sp_loc),
                  data = d, REML = FALSE
                )
                  # # (Covid|IDLocality) +
      est_m1b = est_out(m1b, "01b) (1|Year) + (1|weekday) + (scale(Covid)|genus) + (scale(Covid)|Species) + (1|sp_day_year) + (scale(Covid)|Country) + (scale(Covid)|IDLocality) + (scale(Covid)|sp_loc)")
    # 01c all data, all random slopes, but some without cor to avoid singularity
      m1c=lmer(scale(log(FID))~
                  scale(log(SD))+
                  scale(log(FlockSize))+
                  scale(log(BodyMass))+
                  scale(sin(rad)) + scale(cos(rad)) + 
                  #scale(Day)+
                  scale(Temp)+
                  scale(Covid)+
                  #(1|Year) +(0+Covid|genus)+(0+Covid|Species)+(1|sp_day_year) + (Covid|Country) + (0+Covid|IDLocality) +(Covid|sp_loc)
                  (1|Year)+ (1|weekday) + (0+scale(Covid)|genus)+(0+scale(Covid)|Species)+(1|sp_day_year) + (scale(Covid)|Country) +(scale(Covid)|IDLocality) + (scale(Covid)|sp_loc),
                  data = d, REML = FALSE) # (0+scale(Covid)|genus) is zero 
      est_m1c = est_out(m1c, "01c) (1|Year) + (1|weekday) + (0+scale(Covid)|genus) + (0+scale(Covid)|Species) + (1|sp_day_year) + (scale(Covid)|Country) + (scale(Covid)|IDLocality) + (scale(Covid)|sp_loc)")
    # 01d all data, random slopes that allow for non-singular fit 
      m1d=lmer(scale(log(FID))~
                  scale(log(SD))+
                  scale(log(FlockSize))+
                  scale(log(BodyMass))+
                  scale(sin(rad)) + scale(cos(rad)) + 
                  #scale(Day)+
                  scale(Temp)+
                  scale(Covid)+
                  (1|Year) + (1|weekday) +(1|genus)+(1|Species)+(1|sp_day_year) + (scale(Covid)|Country) + (scale(Covid)|IDLocality) + (scale(Covid)|sp_loc),
                  data = d, REML = FALSE)# (Covid|IDLocality) +
      #d[,res := resid(mf)]
      est_m1d = est_out(m1d, "01d) (1|Year) + (1|weekday)+ (1|genus) + (1|Species) + (1|sp_day_year) + (scale(Covid)|Country) + (scale(Covid)|IDLocality) + (scale(Covid)|sp_loc)")
    # 01e all data, random slopes that allow for non-singular fit with simple structure
      m1e=lmer(scale(log(FID))~
                  scale(log(SD))+
                  scale(log(FlockSize))+
                  scale(log(BodyMass))+
                  scale(sin(rad)) + scale(cos(rad)) + 
                  #scale(Day)+
                  scale(Temp)+
                  scale(Covid)+
                  (1|Year) + (1|weekday)+(1|genus)+(1|Species)+(1|sp_day_year) + (scale(Covid)|Country) + (1|IDLocality) + (1|sp_loc),
                  data = d, REML = FALSE)# (Covid|IDLocality) +
      est_m1e = est_out(m1e, "01e) (1|Year) + (1|weekday) + (1|genus) + (1|Species) + (1|sp_day_year) + (scale(Covid)|Country) + (1|IDLocality) + (1|sp_loc)")

    # 02a) before & during > 4/species - main text 
      dx = dd[N_during>4 & N_before >4]
      #dxx = d[Species %in% dx$Species]
      #dx2 = dd[N_during>9 & N_before >9]
      m2a <- lmer(scale(log(FID)) ~
        scale(log(SD)) +
        scale(log(FlockSize)) +
        scale(log(BodyMass)) +
        scale(sin(rad)) + scale(cos(rad)) +
        # scale(Day)+
        scale(Temp) +
        scale(Covid) +
        (1 | Year) + (1| weekday) + (1 | genus) + (1 | Species) + (1 | sp_day_year) + (scale(Covid) | Country) + (scale(Covid) | IDLocality) + (1 | sp_loc),
      data = d[Species %in% dx$Species], REML = FALSE,
       control = lmerControl(
         optimizer = "optimx", optCtrl = list(method = "nlminb")
       )
      )
      est_m2a = est_out(m2a, "02a) (1|Year) + (1|weekday) + (1|genus) + (1|Species) + (1|sp_day_year) + (scale(Covid)|Country) + (scale(Covid)|IDLocality) + (1|sp_loc); >4/species/period")
    
    # 02b) before & during > 4/species - singularity
      dx = dd[N_during > 4 & N_before > 4]
      m2b=lmer(scale(log(FID))~
                  scale(log(SD))+
                  scale(log(FlockSize))+
                  scale(log(BodyMass))+
                  scale(sin(rad)) + scale(cos(rad)) + 
                  #scale(Day)+
                  scale(Temp)+
                  scale(Covid)+
                  (1|Year) + (1| weekday)+ (scale(Covid)|genus)+(scale(Covid)|Species)+(1|sp_day_year) + (scale(Covid)|Country) +(scale(Covid)|IDLocality) + (scale(Covid)|sp_loc),
                  #(1|Year) + (1|genus)+(1|Species)+(1|sp_day_year) + (Covid|Country) + (Covid|sp_loc),
                  data = d[Species %in% dx$Species],
                  REML = FALSE) # (Covid|IDLocality) +
      est_m2b = est_out(m2b, "02b) (1|Year) + (1|weekday)+ (scale(Covid)|genus)+(scale(Covid)|Species) + (1|sp_day_year) + (scale(Covid)|Country) + (scale(Covid)|IDLocality) + (scale(Covid)|sp_loc); >4/species/period")
    # 02c) before & during > 4/species, random slopes that allow for non-singular fit and simple structure
      dx = dd[N_during>4 & N_before >4]
      #dxx = d[Species %in% dx$Species]
      #dx2 = dd[N_during>9 & N_before >9]
      m2c=lmer(scale(log(FID))~
                  scale(log(SD))+
                  scale(log(FlockSize))+
                  scale(log(BodyMass))+
                  scale(sin(rad)) + scale(cos(rad)) + 
                  #scale(Day)+
                  scale(Temp)+
                  scale(Covid)+
                  (1|Year)+ (1| weekday) +(1|genus)+(1|Species)+(1|sp_day_year) + (scale(Covid)|Country) + (1|IDLocality) + (1|sp_loc),
                  #(1|Year) + (1|genus)+(1|Species)+(1|sp_day_year) + (Covid|Country) + (Covid|sp_loc),
                  data = d[Species %in% dx$Species],
                  REML = FALSE) # (Covid|IDLocality) +
      est_m2c = est_out(m2c, "02c) (1|Year) + (1|weekday)+ (1|genus) + (1|Species)+(1|sp_day_year) + (scale(Covid)|Country) + (1|IDLocality) + (1|sp_loc); >4/species/period")
    # 03a) before & during > 9/species - main text
      dx = dd[N_during > 9 & N_before > 9]
      m3a <- lmer(scale(log(FID)) ~
       scale(log(SD)) +
       scale(log(FlockSize)) +
       scale(log(BodyMass)) +
       scale(sin(rad)) + scale(cos(rad)) +
       # scale(Day)+
       scale(Temp) +
       scale(Covid) +
       (1 | Year) + (1 | weekday) + (1 | genus) + (1 | Species) + (1 | sp_day_year) + (scale(Covid) | Country) + (scale(Covid) | IDLocality) + (1 | sp_loc),
      data = d[Species %in% dx$Species], REML = FALSE,
      )
     est_m3a = est_out(m3a, "03a) (1|Year) + (1|weekday) + (1|genus) + (1|Species) + (1|sp_day_year) + (scale(Covid)|Country) + (scale(Covid)|IDLocality) + (1|sp_loc); >9/species/period")
    # 03b) before & during > 9/species - singularity 
      dx = dd[N_during>9 & N_before >9]
      #dxx = d[Species %in% dx$Species]
      #dx2 = dd[N_during>9 & N_before >9]
      m3b=lmer(scale(log(FID))~
                  scale(log(SD))+
                  scale(log(FlockSize))+
                  scale(log(BodyMass))+
                  scale(sin(rad)) + scale(cos(rad)) + 
                  #scale(Day)+
                  scale(Temp)+
                  scale(Covid)+
                  (1|Year) + (1| weekday) + (scale(Covid)|genus)+(scale(Covid)|Species)+(1|sp_day_year) + (scale(Covid)|Country) +(scale(Covid)|IDLocality) + (scale(Covid)|sp_loc),
                  #(1|Year) + (1|genus)+(1|Species)+(1|sp_day_year) + (Covid|Country) + (Covid|sp_loc),
                  data = d[Species %in% dx$Species],
                  REML = FALSE) # (Covid|IDLocality) +
      est_m3b = est_out(m3b, "03b) (1|Year) + (1|weekday) + (scale(Covid)|genus)+(scale(Covid)|Species) + (1|sp_day_year) + (scale(Covid)|Country) + (scale(Covid)|IDLocality) + (scale(Covid)|sp_loc); >9/species/period")
    # 03c) before & during > 9/species, random slopes that allow for non-singular fit and simple structure
      dx = dd[N_during>9 & N_before >9]
      #dxx = d[Species %in% dx$Species]
      #dx2 = dd[N_during>9 & N_before >9]
      m3c=lmer(scale(log(FID))~
                  scale(log(SD))+
                  scale(log(FlockSize))+
                  scale(log(BodyMass))+
                  scale(sin(rad)) + scale(cos(rad)) + 
                  #scale(Day)+
                  scale(Temp)+
                  scale(Covid)+
                  (1|Year) + (1| weekday) +(1|genus)+(1|Species)+(1|sp_day_year) + (scale(Covid)|Country) + (1|IDLocality) + (1|sp_loc),
                  #(1|Year) + (1|genus)+(1|Species)+(1|sp_day_year) + (Covid|Country) + (Covid|sp_loc),
                  data = d[Species %in% dx$Species],
                  REML = FALSE) # (Covid|IDLocality) +
      est_m3c = est_out(m3c, '03c) (1|Year) + (1|genus) + (1|Species)+(1|sp_day_year) + (scale(Covid)|Country) + (1|IDLocality) + (1|sp_loc); >9/species/period')  
  
  # prepare estimates Stringency 
    m01a <- lmer(scale(log(FID)) ~
      scale(Year) +
      scale(log(SD)) +
      scale(log(FlockSize)) +
      scale(log(BodyMass)) +
      scale(sin(rad)) + scale(cos(rad)) +
      # scale(Day)+
      scale(Temp) +
      scale(StringencyIndex) +
      (1|weekday) + (1 | genus) + (1 | Species) + (1 | sp_day_year) + (scale(StringencyIndex) | Country) + (1 | IDLocality) + (1 | sp_loc),
      data = s, REML = FALSE)
    est_m01a = est_out(m01a, "01a) (1|weekday) + (1|genus) + (1|Species) + (1|sp_day_year) + (scale(StringencyIndex)|Country) + (1|IDLocality) +(1|sp_loc)")

    m01b=lmer(scale(log(FID))~ 
        scale(Year)+ 
        scale(log(SD))+ 
        scale(log(FlockSize))+ 
        scale(log(BodyMass))+ 
        scale(sin(rad)) + scale(cos(rad)) +  
        #scale(Day)+ 
        scale(Temp)+ 
        scale(StringencyIndex)+ 
        (1|weekday) + (scale(StringencyIndex)|genus)+(1|Species)+(1|sp_day_year) + (scale(StringencyIndex)|Country) + (1|IDLocality) +(1|sp_loc),
        data = s, REML = FALSE
        )  
        # (1|Year) explains nothing - could stay 
     est_m01b = est_out(m01b, "01b) (1|weekday) + (scale(StringencyIndex)|genus) + (1|Species) + (1|sp_day_year) + (scale(StringencyIndex)|Country) + (1|IDLocality) +(1|sp_loc)")
     
     m01c=lmer(scale(log(FID))~
          scale(Year)+       
          scale(log(SD))+
          scale(log(FlockSize))+
          scale(log(BodyMass))+
          scale(sin(rad)) + scale(cos(rad)) + 
          #scale(Day)+
          scale(Temp)+
          scale(StringencyIndex)+
          (scale(StringencyIndex)|Country) + (1|IDLocality),  
          data = s, REML = FALSE) 
          # (1|Year), (1|genus), (1|sp_day_year), (1|sp_loc),  explain nothing - could stay
     est_m01c = est_out(m01c, "01c) (scale(StringencyIndex)|Country) + (1|IDLocality)")

     m02a <- lmer(scale(log(FID)) ~
       scale(Year) +
       scale(log(SD)) +
       scale(log(FlockSize)) +
       scale(log(BodyMass)) +
       scale(sin(rad)) + scale(cos(rad)) +
       # scale(Day)+
       scale(Temp) +
       scale(StringencyIndex) +
       (1 | weekday) + (1 | genus) + (1 | Species) + (1 | sp_day_year) + (scale(StringencyIndex) | Country) + (1 | IDLocality) + (1 | sp_loc),
     data = s[Nsp>4], REML = FALSE
     )

     est_m02a = est_out(m02a, "02a) (1|weekday) + (1|genus) + (1|Species) + (1|sp_day_year) + (scale(StringencyIndex)|Country) + (1|IDLocality) +(1|sp_loc); >4/species")
    m02b=lmer(scale(log(FID))~ 
        scale(Year)+ 
        scale(log(SD))+ 
        scale(log(FlockSize))+ 
        scale(log(BodyMass))+ 
        scale(sin(rad)) + scale(cos(rad)) +  
        #scale(Day)+ 
        scale(Temp)+ 
        scale(StringencyIndex)+ 
        (1|weekday) + (scale(StringencyIndex)|genus)+(1|Species)+(1|sp_day_year) + (scale(StringencyIndex)|Country) + (1|IDLocality) +(1|sp_loc),
        data = s[Nsp>4], REML = FALSE,  
        control = lmerControl( 
            optimizer ='optimx', optCtrl=list(method='nlminb')) 
        )  
        # (1|Year) explains nothing - could stay 
     est_m02b = est_out(m02b, "02b) (1|weekday) + (scale(StringencyIndex)|genus) + (1|Species) + (1|sp_day_year) + (scale(StringencyIndex)|Country) + (1|IDLocality) +(1|sp_loc); >4/species")
     
     m02c=lmer(scale(log(FID))~
          scale(Year)+       
          scale(log(SD))+
          scale(log(FlockSize))+
          scale(log(BodyMass))+
          scale(sin(rad)) + scale(cos(rad)) + 
          #scale(Day)+
          scale(Temp)+
          scale(StringencyIndex)+
          (scale(StringencyIndex)|Country) + (1|IDLocality),  
          data = s[Nsp>4], REML = FALSE) 
          # (1|Year), (1|genus), (1|sp_day_year), (1|sp_loc),  explain nothing - could stay
      est_m02c = est_out(m02c, "02c) (scale(StringencyIndex)|Country) + (1|IDLocality); >4/species")

      m03a <- lmer(scale(log(FID)) ~
           scale(Year) +
           scale(log(SD)) +
           scale(log(FlockSize)) +
           scale(log(BodyMass)) +
           scale(sin(rad)) + scale(cos(rad)) +
           # scale(Day)+
           scale(Temp) +
           scale(StringencyIndex) +
           (1 | weekday) + (1 | genus) + (1 | Species) + (1 | sp_day_year) + (scale(StringencyIndex) | Country) + (1 | IDLocality) + (1 | sp_loc),
         data = s[Nsp > 9], REML = FALSE
         )

         est_m03a = est_out(m03a, "03a) (1|weekday) + (1|genus) + (1|Species) + (1|sp_day_year) + (scale(StringencyIndex)|Country) + (1|IDLocality) +(1|sp_loc); >9/species")
         m03b = lmer(scale(log(FID)) ~
           scale(Year) +
           scale(log(SD)) +
           scale(log(FlockSize)) +
           scale(log(BodyMass)) +
           scale(sin(rad)) + scale(cos(rad)) +
           # scale(Day)+
           scale(Temp) +
           scale(StringencyIndex) +
           (1 | weekday) + (scale(StringencyIndex) | genus) + (1 | Species) + (1 | sp_day_year) + (scale(StringencyIndex) | Country) + (1 | IDLocality) + (1 | sp_loc),
         data = s[Nsp > 9], REML = FALSE
         )
         est_m03b = est_out(m03b, "03b) (1|weekday) + (scale(StringencyIndex)|genus) + (1|Species) + (1|sp_day_year) + (scale(StringencyIndex)|Country) + (1|IDLocality) +(1|sp_loc); >9/species")

        m03c = lmer(scale(log(FID)) ~
           scale(Year) +
           scale(log(SD)) +
           scale(log(FlockSize)) +
           scale(log(BodyMass)) +
           scale(sin(rad)) + scale(cos(rad)) +
           # scale(Day)+
           scale(Temp) +
           scale(StringencyIndex) +
           (scale(StringencyIndex) | Country) + (1 | IDLocality),
         data = s[Nsp > 9], REML = FALSE
         )
         # (1|Year), (1|genus), (1|sp_day_year), (1|sp_loc),  explain nothing - could stay
         est_m03c = est_out(m03c, "03c) (scale(StringencyIndex)|Country) + 1|IDLocality); >9/species")

  # prepare estimates google mobility 
    mg01a=lmer(scale(log(FID))~
      scale(Year)+       
      scale(log(SD))+
      scale(log(FlockSize))+
      scale(log(BodyMass))+
      scale(sin(rad)) + scale(cos(rad)) + 
      #scale(Day)+
      scale(Temp)+
      scale(parks_percent_change_from_baseline)+
      (1|weekday) + (1|genus) +(1|Species) + (1|sp_day_year) + (scale(parks_percent_change_from_baseline)|Country) + (1|IDLocality)+(1|sp_loc),  
      data = ss, REML = FALSE
      ) 
      # (1|Year), (1|genus), (1|sp_day_year), (1|sp_loc),  explain nothing - could stay
    est_mg01a = est_out(mg01a, "01a) (1|weekday) + (1|genus) +(1|Species) + (1|sp_day_year) + (scale(parks_percent_change_from_baseline)|Country) + (1|IDLocality)+(1|sp_loc)")
    
    mg01b <- lmer(scale(log(FID)) ~
     scale(Year) +
     scale(log(SD)) +
     scale(log(FlockSize)) +
     scale(log(BodyMass)) +
     scale(sin(rad)) + scale(cos(rad)) +
     # scale(Day)+
     scale(Temp) +
     scale(parks_percent_change_from_baseline) +
     (1|weekday) + (scale(parks_percent_change_from_baseline)| genus) + (1 | Species) + (1 | sp_day_year) + 
     (scale(parks_percent_change_from_baseline)| Country) + (1 | IDLocality) + (1 | sp_loc),
     data = ss, REML = FALSE
     ) 

     est_mg01b = est_out(mg01b, "01b) (1|weeekday) + (scale(parks_percent_change_from_baseline)|genus)+(1|Species)+(1|sp_day_year) + (scale(parks_percent_change_from_baseline)|Country) + (1|IDLocality) +(1|sp_loc)")
     
    mg01c <- lmer(scale(log(FID)) ~
      scale(Year) +
      scale(log(SD)) +
      scale(log(FlockSize)) +
      scale(log(BodyMass)) +
      scale(sin(rad)) + scale(cos(rad)) +
      # scale(Day)+
      scale(Temp) +
      scale(parks_percent_change_from_baseline) +
      (scale(parks_percent_change_from_baseline) | Country) + (1 | IDLocality),
    data = ss, REML = FALSE
    )
    est_mg01c = est_out(mg01c, "01c) (scale(parks_percent_change_from_baseline)|Country) + (1|IDLocality)")

    mg02a = lmer(scale(log(FID)) ~
      scale(Year) +
      scale(log(SD)) +
      scale(log(FlockSize)) +
      scale(log(BodyMass)) +
      scale(sin(rad)) + scale(cos(rad)) +
      # scale(Day)+
      scale(Temp) +
      scale(parks_percent_change_from_baseline) +
      (1 | weekday) + (1 | genus) + (1 | Species) + (1 | sp_day_year) + (scale(parks_percent_change_from_baseline) | Country) + (1 | IDLocality) + (1 | sp_loc),
    data = ss[Nsp>4], REML = FALSE
    )
    # (1|Year), (1|genus), (1|sp_day_year), (1|sp_loc),  explain nothing - could stay
    est_mg02a = est_out(mg02a, "02a) (1|weekday) + (1|genus) +(1|Species) + (1|sp_day_year) + (scale(parks_percent_change_from_baseline)|Country) + (1|IDLocality)+(1|sp_loc); >4/specie")

    mg02b <- lmer(scale(log(FID)) ~
      scale(Year) +
      scale(log(SD)) +
      scale(log(FlockSize)) +
      scale(log(BodyMass)) +
      scale(sin(rad)) + scale(cos(rad)) +
      # scale(Day)+
      scale(Temp) +
      scale(parks_percent_change_from_baseline) +
      (1 | weekday) + (scale(parks_percent_change_from_baseline) | genus) + (1 | Species) + (1 | sp_day_year) +
      (scale(parks_percent_change_from_baseline) | Country) + (1 | IDLocality) + (1 | sp_loc),
    data = ss[Nsp>4], REML = FALSE
    )

    est_mg02b = est_out(mg02b, "02b) (1|weeekday) + (scale(parks_percent_change_from_baseline)|genus)+(1|Species)+(1|sp_day_year) + (scale(parks_percent_change_from_baseline)|Country) + (1|IDLocality) +(1|sp_loc); >4/specie")

    mg02c <- lmer(scale(log(FID)) ~
      scale(Year) +
      scale(log(SD)) +
      scale(log(FlockSize)) +
      scale(log(BodyMass)) +
      scale(sin(rad)) + scale(cos(rad)) +
      # scale(Day)+
      scale(Temp) +
      scale(parks_percent_change_from_baseline) +
      (scale(parks_percent_change_from_baseline) | Country) + (1 | IDLocality),
    data = ss[Nsp>4], REML = FALSE
    )
    est_mg02c = est_out(mg02c, "02c) (scale(parks_percent_change_from_baseline)|Country) + (1|IDLocality); >4/specie")

    mg03a = lmer(scale(log(FID)) ~
      scale(Year) +
      scale(log(SD)) +
      scale(log(FlockSize)) +
      scale(log(BodyMass)) +
      scale(sin(rad)) + scale(cos(rad)) +
      # scale(Day)+
      scale(Temp) +
      scale(parks_percent_change_from_baseline) +
      (1 | weekday) + (1 | genus) + (1 | Species) + (1 | sp_day_year) + (scale(parks_percent_change_from_baseline) | Country) + (1 | IDLocality) + (1 | sp_loc),
    data = ss[Nsp > 9], REML = FALSE
    )
    # (1|Year), (1|genus), (1|sp_day_year), (1|sp_loc),  explain nothing - could stay
    est_mg03a = est_out(mg03a, "03a) (1|weekday) + (1|genus) +(1|Species) + (1|sp_day_year) + (scale(parks_percent_change_from_baseline)|Country) + (1|IDLocality)+(1|sp_loc); >9/specie")

    mg03b <- lmer(scale(log(FID)) ~
      scale(Year) +
      scale(log(SD)) +
      scale(log(FlockSize)) +
      scale(log(BodyMass)) +
      scale(sin(rad)) + scale(cos(rad)) +
      # scale(Day)+
      scale(Temp) +
      scale(parks_percent_change_from_baseline) +
      (1 | weekday) + (scale(parks_percent_change_from_baseline) | genus) + (1 | Species) + (1 | sp_day_year) +
      (scale(parks_percent_change_from_baseline) | Country) + (1 | IDLocality) + (1 | sp_loc),
    data = ss[Nsp > 9], REML = FALSE
    )

    est_mg03b = est_out(mg03b, "03b) (1|weeekday) + (scale(parks_percent_change_from_baseline)|genus)+(1|Species)+(1|sp_day_year) + (scale(parks_percent_change_from_baseline)|Country) + (1|IDLocality) +(1|sp_loc); >9/specie")

    mg03c <- lmer(scale(log(FID)) ~
      scale(Year) +
      scale(log(SD)) +
      scale(log(FlockSize)) +
      scale(log(BodyMass)) +
      scale(sin(rad)) + scale(cos(rad)) +
      # scale(Day)+
      scale(Temp) +
      scale(parks_percent_change_from_baseline) +
      (scale(parks_percent_change_from_baseline) | Country) + (1 | IDLocality),
    data = ss[Nsp > 9], REML = FALSE
    )
    est_mg03c = est_out(mg03c, "03c) (scale(parks_percent_change_from_baseline)|Country) + (1|IDLocality); >9/specie")

    # export
      save(file = "Data/Fig_S2_estimates.Rdata", 
      est_m1a, est_m1b, est_m1c, est_m1d, est_m1e, est_m2a, est_m2b, est_m2c, est_m3a, est_m3b, est_m3c, 
      est_m01a, est_m01b, est_m01c, est_m02a, est_m02b, est_m02c, est_m03a, est_m03b, est_m03c, 
      est_mg01a, est_mg01b, est_mg01c, est_mg02a, est_mg02b, est_mg02c, est_mg03a, est_mg03b, est_mg03c)
     
  # Figure S2
     # prepare plot for Period
     xs = rbind(est_m1a, est_m1b, est_m1c, est_m1d, est_m1e, est_m2a, est_m2b, est_m2c, est_m3a, est_m3b, est_m3c)
     xs[, model := gsub("Covid", "Period", xs$model)]
     xs[, model := gsub("Year", "year", xs$model)]
     xs[, model := gsub("Species", "species", xs$model)]
     xs[, model := gsub("sp_day_year", "species within day & year", xs$model)]
     xs[, model := gsub("IDLocality", "site", xs$model)]
     xs[, model := gsub("sp_loc", "species within site", xs$model)]

     gs2 =
       ggplot(xs[predictor == "scale(Covid)"], aes(y = model, x = estimate, col = model)) +
       geom_vline(xintercept = 0, col = "grey30", lty = 3) +
       geom_errorbar(aes(xmin = lwr, xmax = upr, col = model), width = 0, position = position_dodge(width = 0.01)) +
       # ggtitle ("Sim based")+
       geom_point(position = position_dodge(width = 0.01)) +
       # scale_colour_brewer(type = 'qual', palette = 'Paired',guide = guide_legend(reverse = TRUE))+
       # scale_fill_brewer(type = 'qual', palette = 'Paired',guide = guide_legend(reverse = TRUE))+
       scale_color_viridis(discrete = TRUE, guide = guide_legend(reverse = TRUE)) +
       scale_fill_viridis(discrete = TRUE, guide = guide_legend(reverse = TRUE)) +
       scale_y_discrete(limits = rev) +
       coord_fixed(ratio = 0.05, xlim = c(-0.23, 0.15)) +
       # scale_shape(guide = guide_legend(reverse = TRUE)) +
       # scale_x_continuous(limits = c(-2, 2), expand = c(0, 0), breaks = seq(-2,2, by = 1), labels = seq(-2,2, by = 1)) +
       labs(y = NULL, x = "Period (before vs during shutdowns)", tag = "a)") + # title = "a)",Effect of Period (before/during shutdown)
       # ylim(c(0,100))+
       # coord_flip()+
       theme_bw() +
       theme(
         legend.position = "none",
         #plot.title.position = "plot",
         plot.subtitle = element_text(size = 7),
         plot.title = element_text(size = 7),
         plot.tag = element_text(size = 7),
         legend.title = element_text(size = 7),
         legend.text = element_text(size = 6),
         ## legend.spacing.y = unit(0.1, 'cm'),
         legend.key.height = unit(0.5, "line"),
         # plot.margin = margin(b = 0.5, l = 0.5, t = 0.5, r =0.5, unit =  "pt"),
         panel.grid = element_blank(),
         panel.border = element_blank(),
         panel.background = element_blank(),
         axis.line = element_line(colour = ax_lines, size = 0.25),
         axis.line.y = element_blank(),
         axis.ticks.y = element_blank(),
         axis.ticks.x = element_line(colour = ax_lines, size = 0.25),
         axis.ticks.length = unit(1, "pt"),
         axis.text.x = element_text(colour = "black", size = 6),
         axis.text.y = element_text(colour = "black", size = 7),
         axis.title = element_text(size = 7)
       )
     #gs2
     # ggsave(here::here('Outputs/Figure_Sy.png'),g, width = 30, height =5, units = 'cm')
     # prepare plot for Stringency
     xs0 = rbind(est_m01a, est_m01b, est_m01c, est_m02a, est_m02b, est_m02c, est_m03a, est_m03b, est_m03c)
     xs0[, model := gsub("StringencyIndex", "Stringency Index", model)]
     xs0[, model := gsub("Year", "year", model)]
     xs0[, model := gsub("Species", "species", model)]
     xs0[, model := gsub("sp_day_year", "species within day & year", model)]
     xs0[, model := gsub("IDLocality", "site", model)]
     xs0[, model := gsub("sp_loc", "species within site", model)]
     g0 =
       ggplot(xs0[predictor == "scale(StringencyIndex)"], aes(y = model, x = estimate, col = model)) +
       geom_vline(xintercept = 0, col = "grey30", lty = 3) +
       geom_errorbar(aes(xmin = lwr, xmax = upr, col = model), width = 0, position = position_dodge(width = 0.01)) +
       # ggtitle ("Sim based")+
       geom_point(position = position_dodge(width = 0.01)) +
       # scale_colour_brewer(type = 'qual', palette = 'Paired',guide = guide_legend(reverse = TRUE))+
       # scale_fill_brewer(type = 'qual', palette = 'Paired',guide = guide_legend(reverse = TRUE))+
       scale_color_viridis(discrete = TRUE, guide = guide_legend(reverse = TRUE)) +
       scale_fill_viridis(discrete = TRUE, guide = guide_legend(reverse = TRUE)) +
       scale_y_discrete(limits = rev) +
       coord_fixed(ratio = 0.05, xlim = c(-0.23, 0.15)) +
       # scale_shape(guide = guide_legend(reverse = TRUE)) +
       # scale_x_continuous(limits = c(-2, 2), expand = c(0, 0), breaks = seq(-2,2, by = 1), labels = seq(-2,2, by = 1)) +
       labs(y = NULL, x = "Stringency index", tag = "b)") + # title = "b) Effect of ") +
       # ylim(c(0,100))+
       # coord_flip()+
       theme_bw() +
       theme(
         legend.position = "none",
         plot.subtitle = element_text(size = 7),
         plot.title = element_text(size = 7),
         plot.tag = element_text(size = 7),
         legend.title = element_text(size = 7),
         legend.text = element_text(size = 6),
         ## legend.spacing.y = unit(0.1, 'cm'),
         legend.key.height = unit(0.5, "line"),
         # plot.margin = margin(b = 0.5, l = 0.5, t = 0.5, r =0.5, unit =  "pt"),
         panel.grid = element_blank(),
         panel.border = element_blank(),
         panel.background = element_blank(),
         axis.line = element_line(colour = ax_lines, size = 0.25),
         axis.line.y = element_blank(),
         axis.ticks.y = element_blank(),
         axis.ticks.x = element_line(colour = ax_lines, size = 0.25),
         axis.ticks.length = unit(1, "pt"),
         axis.text.x = element_text(colour = "black", size = 6),
         axis.text.y = element_text(colour = "black", size = 7),
         axis.title = element_text(size = 7)
       )
     #g0
     # ggsave(here::here('Outputs/Figure_Sz.png'),g0, width = 30, height =5, units = 'cm')
     # prepare plot for Google
     xg0 = rbind(est_mg01a, est_mg01b, est_mg01c, est_mg02a, est_mg02b, est_mg02c, est_mg03a, est_mg03b, est_mg03c)
     xg0[, model := gsub("parks_percent_change_from_baseline", "Google Mobility", model)]
     xg0[, model := gsub("Year", "year", model)]
     xg0[, model := gsub("Species", "species", model)]
     xg0[, model := gsub("sp_day_year", "species within day & year", model)]
     xg0[, model := gsub("IDLocality", "site", model)]
     xg0[, model := gsub("sp_loc", "species within site", model)]
     gg0 =
       ggplot(xg0[predictor == "scale(parks_percent_change_from_baseline)"], aes(y = model, x = estimate, col = model)) +
       geom_vline(xintercept = 0, col = "grey30", lty = 3) +
       geom_errorbar(aes(xmin = lwr, xmax = upr, col = model), width = 0, position = position_dodge(width = 0.01)) +
       # ggtitle ("Sim based")+
       geom_point(position = position_dodge(width = 0.01)) +
       # scale_colour_brewer(type = 'qual', palette = 'Paired',guide = guide_legend(reverse = TRUE))+
       # scale_fill_brewer(type = 'qual', palette = 'Paired',guide = guide_legend(reverse = TRUE))+
       scale_color_viridis(discrete = TRUE, guide = guide_legend(reverse = TRUE)) +
       scale_fill_viridis(discrete = TRUE, guide = guide_legend(reverse = TRUE)) +
       scale_y_discrete(limits = rev) +
       coord_fixed(ratio = 0.05, xlim = c(-0.23, 0.15)) +
       # scale_shape(guide = guide_legend(reverse = TRUE)) +
       # scale_x_continuous(limits = c(-2, 2), expand = c(0, 0), breaks = seq(-2,2, by = 1), labels = seq(-2,2, by = 1)) +
       labs(y = NULL, x = "Google Mobility\n[Standardized effect sizes on\nflight initiation distances]", tag = "c)") + # title = "b) Effect of ") +
       # ylim(c(0,100))+
       # coord_flip()+
       theme_bw() +
       theme(
         legend.position = "none",
         plot.subtitle = element_text(size = 7),
         plot.title = element_text(size = 7),
         plot.tag = element_text(size = 7),
         legend.title = element_text(size = 7),
         legend.text = element_text(size = 6),
         ## legend.spacing.y = unit(0.1, 'cm'),
         legend.key.height = unit(0.5, "line"),
         # plot.margin = margin(b = 0.5, l = 0.5, t = 0.5, r =0.5, unit =  "pt"),
         panel.grid = element_blank(),
         panel.border = element_blank(),
         panel.background = element_blank(),
         axis.line = element_line(colour = ax_lines, size = 0.25),
         axis.line.y = element_blank(),
         axis.ticks.y = element_blank(),
         axis.ticks.x = element_line(colour = ax_lines, size = 0.25),
         axis.ticks.length = unit(1, "pt"),
         axis.text.x = element_text(colour = "black", size = 6),
         axis.text.y = element_text(colour = "black", size = 7),
         axis.title = element_text(size = 7)
       )
     #gg0
     # ggsave(here::here('Outputs/Figure_Sz.png'),g0, width = 30, height =5, units = 'cm')
     # combine
     ggsave(here::here("Outputs/Fig_S2_rev_v6.png"), rbind(ggplotGrob(gs2), ggplotGrob(g0), ggplotGrob(gg0)), width = 30, height = 15, units = "cm")
```

## Supplementary TABLES

```{r }
  # prepare effects on FID general
    # PERIOD
     m1a_ = m_out(name = "Table S1 - 1a", dep = "Escape distance", model = m1a, nsim = 5000)
     m1b_ = m_out(name = "Table S1 - 1b", dep = "Escape distance", model = m1b, nsim = 5000)
     m1c_ = m_out(name = "Table S1 - 1c", dep = "Escape distance", model = m1c, nsim = 5000)
     m1d_ = m_out(name = "Table S1 - 1d", dep = "Escape distance", model = m1d, nsim = 5000)
     m1e_ = m_out(name = "Table S1 - 1e", dep = "Escape distance", model = m1e, nsim = 5000)

     m2a_ = m_out(name = "Table S1 - 2a", dep = "Escape distance", model = m2a, nsim = 5000)
     m2b_ = m_out(name = "Table S1 - 2b", dep = "Escape distance", model = m2b, nsim = 5000)
     m2c_ = m_out(name = "Table S1 - 2c", dep = "Escape distance", model = m2c, nsim = 5000)
     m3a_ = m_out(name = "Table S1 - 3a", dep = "Escape distance", model = m3a, nsim = 5000)
     m3b_ = m_out(name = "Table S1 - 3b", dep = "Escape distance", model = m3b, nsim = 5000)
     m3c_ = m_out(name = "Table S1 - 3c", dep = "Escape distance", model = m3c, nsim = 5000)

    # Stringency
     m01a_ = m_out(name = "Table S2 - 1a", dep = "Escape distance", model = m01a, nsim = 5000)
     m01b_ = m_out(name = "Table S2 - 1b", dep = "Escape distance", model = m01b, nsim = 5000)
     m01c_ = m_out(name = "Table S2 - 1c", dep = "Escape distance", model = m01c, nsim = 5000)

     m02a_ = m_out(name = "Table S2 - 2a", dep = "Escape distance", model = m02a, nsim = 5000)
     m02b_ = m_out(name = "Table S2 - 2b", dep = "Escape distance", model = m02b, nsim = 5000)
     m02c_ = m_out(name = "Table S2 - 2c", dep = "Escape distance", model = m02c, nsim = 5000)

     m03a_ = m_out(name = "Table S2 - 3a", dep = "Escape distance", model = m03a, nsim = 5000)
     m03b_ = m_out(name = "Table S2 - 3b", dep = "Escape distance", model = m03b, nsim = 5000)
     m03c_ = m_out(name = "Table S2 - 3c", dep = "Escape distance", model = m03c, nsim = 5000)
    
    # google
     mg01a_ = m_out(name = "Table S3 - 1a", dep = "Escape distance", model = mg01a, nsim = 5000)
     mg01b_ = m_out(name = "Table S3 - 1b", dep = "Escape distance", model = mg01b, nsim = 5000)
     mg01c_ = m_out(name = "Table S3 - 1c", dep = "Escape distance", model = mg01c, nsim = 5000)

     mg02a_ = m_out(name = "Table S3 - 2a", dep = "Escape distance", model = mg02a, nsim = 5000)
     mg02b_ = m_out(name = "Table S3 - 2b", dep = "Escape distance", model = mg02b, nsim = 5000)
     mg02c_ = m_out(name = "Table S3 - 2c", dep = "Escape distance", model = mg02c, nsim = 5000)

     mg03a_ = m_out(name = "Table S3 - 3a", dep = "Escape distance", model = mg03a, nsim = 5000)
     mg03b_ = m_out(name = "Table S3 - 3b", dep = "Escape distance", model = mg03b, nsim = 5000)
     mg03c_ = m_out(name = "Table S3 - 3c", dep = "Escape distancey", model = mg03c, nsim = 5000)

     out1 = rbind(m1a_, m1b_, m1c_, m1d_, m1e_, m2a_, m2b_, m2c_,m3a_, m3b_, m3c_, fill = TRUE)
     out1[is.na(out1)] = ""
       out1[, effect := gsub("Covid", "Period", effect)]
       out1[, effect := gsub("Year", "year", effect)]
       out1[, effect := gsub("Species", "species", effect)]
       out1[, effect := gsub("sp_day_year", "species within day & year", effect)]
       out1[, effect := gsub("IDLocality", "site", effect)]
       out1[, effect := gsub("sp_loc", "species within site", effect)]
     fwrite(file = "./Outputs/Table_S1_rev.csv", out1)

     out2 = rbind(m01a_, m01b_, m01c_, m02a_, m02b_, m02c_, m03a_, m03b_, m03c_, fill = TRUE)
     out2[is.na(out2)] = ""
       out2[, effect := gsub("StringencyIndex", "Stringency Index", effect)]
       out2[, effect := gsub("Year", "year", effect)]
       out2[, effect := gsub("Species", "species", effect)]
       out2[, effect := gsub("sp_day_year", "species within day & year", effect)]
       out2[, effect := gsub("IDLocality", "site", effect)]
       out2[, effect := gsub("sp_loc", "species within site", effect)]
     fwrite(file = "./Outputs/Table_S2_rev.csv", out2)

     out3 = rbind(mg01a_, mg01b_, mg01c_, mg02a_, mg02b_, mg02c_, mg03a_, mg03b_, mg03c_, fill = TRUE)
     out3[is.na(out3)] = ""
       out2[, effect := gsub("parks_percent_change_from_baseline", "Google Mobility", effect)]
       out2[, effect := gsub("Year", "year", effect)]
       out2[, effect := gsub("Species", "species", effect)]
       out2[, effect := gsub("sp_day_year", "species within day & year", effect)]
       out2[, effect := gsub("IDLocality", "site", effect)]
       out2[, effect := gsub("sp_loc", "species within site", effect)]
     fwrite(file = "./Outputs/Table_S3_rev.csv", out3)
  
  # prepare FID - PERIOD country specific 
    ms_out = m_out(name = "Table S_FID_country - full a", dep = "Escape distance", model = ms, nsim = 5000)

    mx_out = m_out(name = "Table S_FID_country - full b", dep = "Escape distance", model = mx, nsim = 5000)

    cs_out = m_out(name = "Table S_FID_country - CZ a", dep = "Escape distancey", model = cs, nsim = 5000)

    cx_out = m_out(name = "Table S_FID_country - CZ b", dep = "Escape distance", model = cx, nsim = 5000)

    fs_out = m_out(name = "Table S_FID_country - FI a", dep = "Escape distance", model = fs, nsim = 5000)

    fx_out =  m_out(name = "Table S_FID_country - FI b", dep = "Escape distance", model = fx, nsim = 5000)

    hs_out = m_out(name = "Table S_FID_country - HU a", dep = "Escape distance", model = hs, nsim = 5000)

    hx_out = m_out(name = "Table S_FID_country - HU b", dep = "Escape distance", model = hx, nsim = 5000)

    as_out = m_out(name = "Table S_FID_country - AU a", dep = "Escape distance", model = as, nsim = 5000)

    ax_out = m_out(name = "Table S_FID_country - AU b", dep = "Escape distance", model = ax, nsim = 5000)

    ps_out = m_out(name = "Table S_FID_country - PL a", dep = "Escape distance", model = ps, nsim = 5000)

    px_out = m_out(name = "Table S_FID_country - PL b", dep = "Escape distance", model = px, nsim = 5000)

    out_FID_c = rbind(fs_out, fx_out, ps_out, px_out, cs_out, cx_out, hs_out,hx_out,as_out, ax_out,  fill = TRUE)

    out_FID_c[is.na(out_FID_c)] = ""
    out_FID_c[, effect := gsub("Covid", "Period", effect)]
    out_FID_c[, effect := gsub("Year", "year", effect)]
    out_FID_c[, effect := gsub("SD", "starting distance", effect)]
    out_FID_c[, effect := gsub("Temp", "temperaturre", effect)]
    out_FID_c[, effect := gsub("FlockSize", "flock size", effect)]
    out_FID_c[, effect := gsub("BodyMass", "body mass", effect)]
    out_FID_c[, effect := gsub("Species", "species", effect)]
    out_FID_c[, effect := gsub("sp_day_year", "species within day & year", effect)]
    out_FID_c[, effect := gsub("IDLocality", "site", effect)]
    out_FID_c[, effect := gsub("sp_loc", "species within site", effect)]
    fwrite(file = "./Outputs/Table_S_FID_c.csv", out_FID_c)
  
  # prepare Google ~ Stringency
    out_g_s[, effect := gsub("Year", "year", effect)]
    out_g_s[, effect := gsub("StringencyIndex", "stringency index", effect)]
    out_g_s[, effect := gsub("year_weekday", "weekday within year", effect)]
    fwrite(file = "./Outputs/Table_S_G-S.csv", out_g_s)
    
  # prepare FID ~ stringency
    mss_out =  m_out(name = "Table S_FID_s - full a", dep = "Escape distance", model = mss, nsim = 5000)

    msx_out = m_out(name = "Table S_FID_s - full b", dep = "Escape distance", model = msx, nsim = 5000)

    css_out = m_out(name = "Table v - CZ a", dep = "Escape distance", model = css, nsim = 5000)

    csx_out = m_out(name = "Table S_FID_s - CZ b", dep = "Escape distance", model = csx, nsim = 5000)

    fss_out = m_out(name = "Table S_FID_s - FI a", dep = "Escape distance", model = fss, nsim = 5000)

    fsx_out = m_out(name = "Table S_FID_s - FI b", dep = "Escape distance", model = fsx, nsim = 5000)

    hss_out = m_out(name = "Table S_FID_s - HU a", dep = "Escape distance", model = hss, nsim = 5000)

    hsx_out = m_out(name = "Table S_FID_s - HU b", dep = "Escape distance", model = hsx, nsim = 5000)

    ass_out = m_out(name = "Table S_FID_s - AU a", dep = "Escape distance", model = ass, nsim = 5000)

    asx_out = m_out(name = "Table S_FID_s - AU b", dep = "Escape distancey", model = asx, nsim = 5000)

    pss_out = m_out(name = "Table S_FID_s - PL a", dep = "Escape distance", model = pss, nsim = 5000)

    psx_out = m_out(name = "Table S_FID_s - PL b", dep = "Escape distancey", model = psx, nsim = 5000)

    out_FID_s = rbind(mss_out, msx_out, fss_out, fsx_out, pss_out, psx_out, css_out, csx_out, hss_out,hsx_out,ass_out, asx_out,  fill = TRUE)

    out_FID_s[is.na(out_FID_c)] = ""
    out_FID_s[, effect := gsub("StringencyIndex", "Stringency index", effect)]
    out_FID_s[, effect := gsub("Year", "year", effect)]
    out_FID_s[, effect := gsub("SD", "starting distance", effect)]
    out_FID_s[, effect := gsub("Temp", "temperaturre", effect)]
    out_FID_s[, effect := gsub("FlockSize", "flock size", effect)]
    out_FID_s[, effect := gsub("BodyMass", "body mass", effect)]
    out_FID_s[, effect := gsub("Species", "species", effect)]
    out_FID_s[, effect := gsub("sp_day_year", "species within day & year", effect)]
    out_FID_s[, effect := gsub("IDLocality", "site", effect)]
    out_FID_s[, effect := gsub("sp_loc", "species within site", effect)]
    fwrite(file = "./Outputs/Table_S_FID_s_c.csv", out_FID_s)

  # prepare FID ~ google
    mgs_out =  m_out(name = "Table S_FID_g - full a", dep = "Escape distance", model = mgs, nsim = 5000)

    mgx_out = m_out(name = "Table S_FID_g - full b", dep = "Escape distance", model = mgx, nsim = 5000)

    cgs_out = m_out(name = "Table S_FID_g - CZ a", dep = "Escape distance", model = cgs, nsim = 5000)

    cgx_out = m_out(name = "Table S_FID_g - CZ b", dep = "Escape distance", model = cgx, nsim = 5000)

    fgs_out = m_out(name = "Table S_FID_g - FI a", dep = "Escape distance", model = fgs, nsim = 5000)

    fgx_out = m_out(name = "Table S_FID_g - FI b", dep = "Escape distance", model = fgx, nsim = 5000)

    hgs_out = m_out(name = "Table S_FID_g - HU a", dep = "Escape distance", model = hgs, nsim = 5000)

    hgx_out = m_out(name = "Table S_FID_g - HU b", dep = "Escape distance", model = hgx, nsim = 5000)

    ags_out = m_out(name = "Table S_FID_g - AU a", dep = "Escape distance", model = ags, nsim = 5000)

    agx_out = m_out(name = "Table S_FID_g - AU b", dep = "Escape distancey", model = agx, nsim = 5000)

    pgs_out = m_out(name = "Table S_FID_g - PL a", dep = "Escape distance", model = pgs, nsim = 5000)

    pgx_out = m_out(name = "Table S_FID_g - PL b", dep = "Escape distancey", model = pgx, nsim = 5000)

    out_FID_g = rbind(mgs_out, mgx_out, fgs_out, fgx_out, pgs_out, pgx_out, cgs_out, cgx_out, hgs_out,hgx_out,ags_out, agx_out,  fill = TRUE)

    out_FID_g[is.na(out_FID_c)] = ""
    out_FID_g[, effect := gsub("parks_percent_change_from_baseline", "Google Mobility", effect)]
    out_FID_g[, effect := gsub("Year", "year", effect)]
    out_FID_g[, effect := gsub("SD", "starting distance", effect)]
    out_FID_g[, effect := gsub("Temp", "temperaturre", effect)]
    out_FID_g[, effect := gsub("FlockSize", "flock size", effect)]
    out_FID_g[, effect := gsub("BodyMass", "body mass", effect)]
    out_FID_g[, effect := gsub("Species", "species", effect)]
    out_FID_g[, effect := gsub("sp_day_year", "species within day & year", effect)]
    out_FID_g[, effect := gsub("IDLocality", "site", effect)]
    out_FID_g[, effect := gsub("sp_loc", "species within site", effect)]
    fwrite(file = "./Outputs/Table_S_FID_g_c.csv", out_FID_g)

# TO DO modelAss

# END
```

